---
title: "Format_count_tables"
author: "Stefano Romano"
output: html_document
editor_options: 
  chunk_output_type: console
---
# Before you begin:

These scripts were tailored for the analyses we conducted. They are not meant to be used in other contexts and are reported here as a detailed extension of the method section of our article. Please, use them at your own risk.

If you use any of the material or approaches in these scripts, please, cite:
Romano et al, 2020, *Meta-analysis of the Parkinson’s disease gut microbiome suggests alterations linked to intestinal inflammation*

Thank you!


# Load the needed pakcages

```{r setup,message=FALSE}
list.of.packages <- c("dplyr","devtools")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

if (!requireNamespace("BiocManager", quietly = TRUE)){
    install.packages("BiocManager")}


list.of.bioc <- c("phyloseq")
new.packages <- list.of.bioc[!(list.of.bioc %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)


library(phyloseq)
library(dplyr)
library(devtools)
```

# Load and clean Species count tables 
Samples with < 4500 reads are removed.
Taxa with counts < 5 in < 2.5.% of samples are removed.
A full path to the Lotus output should be provided.

### Pietrucci et al
```{r Pietrucci}
# Load Species Count table
pietr<-read.delim("Species.txt", header = T, stringsAsFactors = F)
row.names(pietr)<-pietr$Species
pietr<-pietr[,-1]

# Load Lotus mapping file
map<-read.delim("Pietrucci_et_al_mapping.txt", header = T, stringsAsFactors = F)
names(map)[1]<-"SampleID"
unique(map$host_phenotype)

# Check that the samples order is the same in count table and mapping file
all(colnames(pietr) == map$SampleID)

# Create phyloseq object
rownames(map)<-map$SampleID
map<-map[,-1]
pietr_phylo <- phyloseq(otu_table(as.matrix(pietr), taxa_are_rows = T), 
                        sample_data(map))
# Remove taxa that have < 5 counts in < 2.5% taxa
wh0 = genefilter_sample(pietr_phylo, filterfun_sample(function(x) x > 5), A=0.025*nsamples(pietr_phylo))
pietr_phylo = prune_taxa(wh0, pietr_phylo)

# Check whether there are taxa in >= 90% samples
df<-as.data.frame(otu_table(pietr_phylo))
df[df>=1]<-1
if(length(rowSums(df)[rowSums(df) >= 0.9*ncol(df)]) >1){
  print("These are taxa in >90% of samples")
  length(rowSums(df)[rowSums(df) >= 0.9*ncol(df)])
}else{stop("Something is wrong here, there are no taxa present in > 90% samples")}

# Look at the number of PD and HC samples
sum(get_variable(pietr_phylo)$host_phenotype == "Parkinson") 
sum(get_variable(pietr_phylo)$host_phenotype == "Control") 

# Remove samples with < 4500 reads
reads<-read.delim("demulti.acceptsPerSample.log", header = T, stringsAsFactors = F)
if(nrow(subset(reads, Read1Accepted < 4500)) == 0){
  saveRDS(pietr_phylo, 
        "Pietrucci_Species.rds")
}else{
  keep<-subset(reads, Read1Accepted > 4500)$SampleID
  pietr_phylo = prune_samples(keep, pietr_phylo)
  if(any(taxa_sums(pietr_phylo) == 0)){
    pietr_phylo <- prune_taxa(taxa_sums(pietr_phylo) > 0, pietr_phylo)
  }
  saveRDS(pietr_phylo, 
        "Pietrucci_Species.rds")
}

```

### Qian et al
```{r Qian}
# Load Species Count table
qian<-read.delim("Species.txt",header = T, stringsAsFactors = F)
row.names(qian)<-qian$Species
qian<-qian[,-1]

# Load Lotus mapping file
map<-read.delim("Qian_et_al_mapping.txt", header = T, stringsAsFactors = F)
names(map)[1]<-"SampleID"
unique(map$status)

# Check that the samples order is the same in count table and mapping file
qian<-qian[,order(colnames(qian))]
map<-map[order(map$SampleID),]
all(colnames(qian) == map$SampleID)

# Create phyloseq object
rownames(map)<-map$SampleID
map<-map[,-1]
qian_phylo <- phyloseq(otu_table(as.matrix(qian), taxa_are_rows = T), 
                        sample_data(map))


# Remove taxa that have < 5 counts in < 2.5% taxa
wh0 = genefilter_sample(qian_phylo, filterfun_sample(function(x) x > 5), A=0.025*nsamples(qian_phylo))
qian_phylo <- prune_taxa(wh0, qian_phylo)

# Check whether there are taxa in >= 90% samples
df<-as.data.frame(otu_table(qian_phylo))
df[df>=1]<-1
if(length(rowSums(df)[rowSums(df) >= 0.9*ncol(df)]) >1){
  print("There are taxa in >90% of samples")
  length(rowSums(df)[rowSums(df) >= 0.9*ncol(df)])
}else{stop("Something is wrong here, there are not taxa present in > 90% samples")}


# Look at the number of PD and HC samples

sum(get_variable(qian_phylo)$status == "Parkinson") 
sum(get_variable(qian_phylo)$status == "Control") 

# Remove samples with < 4500 reads
reads<-read.delim("demulti.acceptsPerSample.log", header = T, stringsAsFactors = F)
if(nrow(subset(reads, Read1Accepted < 4500)) == 0){
  saveRDS(qian_phylo, 
        "Qian_Species.rds")
}else{
  keep<-subset(reads, Read1Accepted > 4500)$SampleID
  qian_phylo = prune_samples(keep, qian_phylo)
  if(any(taxa_sums(qian_phylo) == 0)){
    qian_phylo <- prune_taxa(taxa_sums(qian_phylo) > 0, qian_phylo)
    }
  saveRDS(qian_phylo, 
        "Qian_Species.rds")
}
```

### Hill-Burns et al
```{r Hill_Burns}
# Load Species Count table
hb<-read.delim("Species.txt", header = T, stringsAsFactors = F)
row.names(hb)<-hb$Species
hb<-hb[,-1]

# Load Lotus mapping file
map<-read.delim("Hill_Burns_et_al_mapping.txt", header = T, stringsAsFactors = F)
names(map)[1]<-"SampleID"

# Correct sample names
names(hb)<-sub("^X1", "1", names(hb))

# Check that the samples order is the same in count table and mapping file
all(colnames(hb) == map$SampleID)

# Create phyloseq object
rownames(map)<-map$SampleID
map<-map[,-1]
hb_phylo <- phyloseq(otu_table(as.matrix(hb), taxa_are_rows = T), 
                        sample_data(map))

# Remove taxa that have < 5 counts in < 2.5% taxa
wh0 = genefilter_sample(hb_phylo, filterfun_sample(function(x) x > 5), A=0.025*nsamples(hb_phylo))
hb_phylo <- prune_taxa(wh0, hb_phylo)

# Check whether there are taxa in >= 90% samples
df<-as.data.frame(otu_table(hb_phylo))
df[df>=1]<-1
if(length(rowSums(df)[rowSums(df) >= 0.9*ncol(df)]) >1){
  print("There are taxa in >90% of samples")
  length(rowSums(df)[rowSums(df) >= 0.9*ncol(df)])
}else{stop("Something is wrong here, there are not taxa present in > 90% samples")}


# Look at hte number of PD and HC samples

sum(get_variable(hb_phylo)$Prkinson == "Parkinson") 
sum(get_variable(hb_phylo)$Prkinson == "Control") 

# Remove samples with < 4500 reads
reads<-read.delim("demulti.acceptsPerSample.log", header = T, stringsAsFactors = F)

if(nrow(subset(reads, ReadsAccepted < 4500)) == 0){
  saveRDS(hb_phylo, 
        "Hill_Burns_Species.rds")
}else{
  keep<-subset(reads, ReadsAccepted > 4500)$SampleID
  hb_phylo = prune_samples(keep, hb_phylo)
  if(any(taxa_sums(hb_phylo) == 0)){
    hb_phylo <- prune_taxa(taxa_sums(hb_phylo) > 0, hb_phylo)
    }
  saveRDS(hb_phylo, 
        "Hill_Burns_Species.rds")
}

```

### Heinz-Bushart et al
```{r Heinz-Bushart }
# Load Species Count table
heinz<-read.delim("Species.txt", header = T, stringsAsFactors = F)
row.names(heinz)<-heinz$Species
heinz<-heinz[,-1]

# Load Lotus mapping file
map<-read.delim("Heintz_Buschart_et_al_mapping.txt", header = T, stringsAsFactors = F)
names(map)[1]<-"SampleID"

# Check that the samples order is the same in count table and mapping file
all(colnames(heinz) == map$SampleID)

# Create phyloseq object
rownames(map)<-map$SampleID
map<-map[,-1]
heinz_phylo <- phyloseq(otu_table(as.matrix(heinz), taxa_are_rows = T), 
                        sample_data(map))

# Remove taxa that have < 5 counts in < 2.5% taxa
wh0 = genefilter_sample(heinz_phylo, filterfun_sample(function(x) x > 5), A=0.025*nsamples(heinz_phylo))
heinz_phylo <- prune_taxa(wh0, heinz_phylo)

# Check whether there are taxa in >= 90% samples
df<-as.data.frame(otu_table(heinz_phylo))
df[df>=1]<-1
if(length(rowSums(df)[rowSums(df) >= 0.9*ncol(df)]) >1){
  print("There are taxa in >90% of samples")
  length(rowSums(df)[rowSums(df) >= 0.9*ncol(df)])
}else{stop("Something is wrong here, there are not taxa present in > 90% samples")}


# Look at the number of PD and HC samples

sum(get_variable(heinz_phylo)$Status == "Parkinson") 
sum(get_variable(heinz_phylo)$Status == "Control") 

# Remove samples with < 4500 reads
reads<-read.delim("demulti.acceptsPerSample.log", header = T, stringsAsFactors = F)

if(nrow(subset(reads, Read1Accepted < 4500)) == 0){
  saveRDS(heinz_phylo, 
        "Heintz_Buschart_Species.rds")
}else{
  keep<-subset(reads, Read1Accepted > 4500)$SampleID
  heinz_phylo = prune_samples(keep, heinz_phylo)
  if(any(taxa_sums(heinz_phylo) == 0)){
    heinz_phylo <- prune_taxa(taxa_sums(heinz_phylo) > 0, heinz_phylo)
    }
  saveRDS(heinz_phylo, 
        "Heintz_Buschart_Species.rds")
}

```

### Petrov et al
```{r Petrov}
# Load Species Count table
petr<-read.delim("Species.txt", header = T, stringsAsFactors = F)
row.names(petr)<-petr$Species
petr<-petr[,-1]

# Load Lotus mapping file
map<-read.delim("Petrov_et_al_onlyPD_mapping.txt", header = T, stringsAsFactors = F)
names(map)[1]<-"SampleID"

# Check that the samples order is the same in count table and mapping file
petr<-petr[,order(colnames(petr))]
map<-map[order(map$SampleID),]
all(colnames(petr) == map$SampleID)

# Create phyloseq object
rownames(map)<-map$SampleID
map<-map[,-1]
petr_phylo <- phyloseq(otu_table(as.matrix(petr), taxa_are_rows = T), 
                        sample_data(map))

# Remove taxa that have < 5 counts in < 2.5% taxa
wh0 <- genefilter_sample(petr_phylo, filterfun_sample(function(x) x > 5), A=0.025*nsamples(petr_phylo))
petr_phylo <- prune_taxa(wh0, petr_phylo)

# Check whether there are taxa in >= 90% samples
df<-as.data.frame(otu_table(petr_phylo))
df[df>=1]<-1
if(length(rowSums(df)[rowSums(df) >= 0.9*ncol(df)]) >1){
  print("There are taxa in >90% of samples")
  length(rowSums(df)[rowSums(df) >= 0.9*ncol(df)])
}else{stop("Something is wrong here, there are not taxa present in > 90% samples")}


# Look at the number of PD and HC samples
sum(get_variable(petr_phylo)$diagnosis == "Parkinson") 
sum(get_variable(petr_phylo)$diagnosis == "Control") 

# Remove samples with < 4500 reads
reads<-read.delim("demulti.acceptsPerSample.log", header = T, stringsAsFactors = F)

if(nrow(subset(reads, ReadsAccepted < 4500)) == 0){
  saveRDS(petr_phylo, 
        "Petrov_Species.rds")
}else{
  keep<-subset(reads, ReadsAccepted > 4500)$SampleID
  petr_phylo <-prune_samples(keep, petr_phylo)
  if(any(taxa_sums(petr_phylo) == 0)){
    petr_phylo <- prune_taxa(taxa_sums(petr_phylo) > 0, petr_phylo)
    }
  saveRDS(petr_phylo, 
        "Petrov_Species.rds")
}

```

### Weis et al
```{r Weis}
# Load Species Count table
weis<-read.delim("Species.txt", header = T, stringsAsFactors = F)
row.names(weis)<-weis$Species
weis<-weis[,-1]

# Load Lotus mapping file
map<-read.delim("Weis_et_al_mapping.txt", header = T, stringsAsFactors = F)
names(map)[1]<-"SampleID"

# Check that the samples order is the same in count table and mapping file
weis<-weis[,order(colnames(weis))]
map<-map[order(map$SampleID),]
all(colnames(weis) == map$SampleID)

# Create phyloseq object
rownames(map)<-map$SampleID
map<-map[,-1]
weis_phylo <- phyloseq(otu_table(as.matrix(weis), taxa_are_rows = T), 
                        sample_data(map))

# Remove taxa that have < 5 counts in < 2.5% taxa
wh0 <- genefilter_sample(weis_phylo, filterfun_sample(function(x) x > 5), A=0.025*nsamples(weis_phylo))
weis_phylo <- prune_taxa(wh0, weis_phylo)

# Check whether there are taxa in >= 90% samples
df<-as.data.frame(otu_table(weis_phylo))
df[df>=1]<-1
if(length(rowSums(df)[rowSums(df) >= 0.9*ncol(df)]) >1){
  print("There are taxa in >90% of samples")
  length(rowSums(df)[rowSums(df) >= 0.9*ncol(df)])
}else{stop("Something is wrong here, there are not taxa present in > 90% samples")}

# Remove samples with < 4500 reads
reads<-read.delim("demulti.acceptsPerSample.log", header = T, stringsAsFactors = F)

if(nrow(subset(reads, ReadsAccepted < 4500)) == 0){
  saveRDS(weis_phylo, 
        "Weis_Species.rds")
}else{
  keep<-subset(reads, ReadsAccepted > 4500)$SampleID
  weis_phylo <-prune_samples(keep, weis_phylo)
  if(any(taxa_sums(weis_phylo) == 0)){
    weis_phylo <- prune_taxa(taxa_sums(weis_phylo) > 0, weis_phylo)
    }
  saveRDS(weis_phylo, 
        "Weis_Species.rds")
}
```

### Scheperjans et al and Aho et al
```{r Scheperjans-Aho}
# Load Species Count table
sa<-read.delim("Species.txt", header = T, stringsAsFactors = F)
row.names(sa)<-sa$Species
sa<-sa[,-1]

# Load Lotus mapping file
map<-read.delim("Scheperjans_Aho_et_al_mapping.txt", header = T, stringsAsFactors = F)
names(map)[1]<-"SampleID"

# Check that the samples order is the same in count table and mapping file
sa<-sa[,order(colnames(sa))]
map<-map[order(map$SampleID),]
all(colnames(sa) == map$SampleID)

# Create phyloseq object
rownames(map)<-map$SampleID
map<-map[,-1]
sa_phylo <- phyloseq(otu_table(as.matrix(sa), taxa_are_rows = T), 
                        sample_data(map))

# Remove taxa that have < 5 counts in < 2.5% taxa
wh0 <- genefilter_sample(sa_phylo, filterfun_sample(function(x) x > 5), A=0.025*nsamples(sa_phylo))
sa_phylo <- prune_taxa(wh0, sa_phylo)

# Look at the number of PD and HC samples
sum(get_variable(hop_phylo)$Status == "Parkinson") 
sum(get_variable(hop_phylo)$Status == "Control") 


# Check whether there are taxa in >= 90% samples
df<-as.data.frame(otu_table(sa_phylo))
df[df>=1]<-1
if(length(rowSums(df)[rowSums(df) >= 0.9*ncol(df)]) >1){
  print("There are taxa in >90% of samples")
  length(rowSums(df)[rowSums(df) >= 0.9*ncol(df)])
}else{stop("Something is wrong here, there are not taxa present in > 90% samples")}


# Remove samples with < 4500 reads
reads<-read.delim("demulti.acceptsPerSample.log", header = T, stringsAsFactors = F)

if(nrow(subset(reads, Read1Accepted < 4500)) == 0){
  saveRDS(sa_phylo, 
        "Scheperjans_Aho_Species.rds")
}else{
  keep<-subset(reads, Read1Accepted > 4500)$SampleID
  sa_phylo <-prune_samples(keep, sa_phylo)
  if(any(taxa_sums(sa_phylo) == 0)){
    sa_phylo <- prune_taxa(taxa_sums(sa_phylo) > 0, sa_phylo)
    }
  saveRDS(sa_phylo, 
        "Scheperjans_Aho_Species.rds")
}
```

### Hopfner et al
```{r Hopfner}
# Load Species Count table
hop<-read.delim("Species.txt", header = T, stringsAsFactors = F)
row.names(hop)<-hop$Species
hop<-hop[,-1]

# Load Lotus mapping file
map<-read.delim("Hopfner_et_al_mapping.txt", header = T, stringsAsFactors = F)
names(map)[1]<-"SampleID"

# Check that the samples order is the same in count table and mapping file
hop<-hop[,order(colnames(hop))]
map<-map[order(map$SampleID),]
all(colnames(hop) == map$SampleID)

# Create phyloseq object
rownames(map)<-map$SampleID
map<-map[,-1]
hop_phylo <- phyloseq(otu_table(as.matrix(hop), taxa_are_rows = T), 
                        sample_data(map))

# Remove taxa that have < 5 counts in < 2.5% taxa
wh0 <- genefilter_sample(hop_phylo, filterfun_sample(function(x) x > 5), A=0.025*nsamples(hop_phylo))
hop_phylo <- prune_taxa(wh0, hop_phylo)

# Check whether there are taxa in >= 90% samples
df<-as.data.frame(otu_table(hop_phylo))
df[df>=1]<-1
if(length(rowSums(df)[rowSums(df) >= 0.9*ncol(df)]) >1){
  print("There are taxa in >90% of samples")
  length(rowSums(df)[rowSums(df) >= 0.9*ncol(df)])
}else{stop("Something is wrong here, there are not taxa present in > 90% samples")}

# Look at the number of PD and HC samples
sum(get_variable(hop_phylo)$Status == "Parkinson") 
sum(get_variable(hop_phylo)$Status == "Control") 

# Remove samples with < 4500 reads
reads<-read.delim("demulti.acceptsPerSample.log", header = T, stringsAsFactors = F)

if(nrow(subset(reads, Read1Accepted < 4500)) == 0){
  saveRDS(hop_phylo, 
        "Hopfner_Species.rds")
}else{
  keep<-subset(reads, Read1Accepted > 4500)$SampleID
  hop_phylo <-prune_samples(keep, hop_phylo)
  if(any(taxa_sums(hop_phylo) == 0)){
    hop_phylo <- prune_taxa(taxa_sums(hop_phylo) > 0, hop_phylo)
    }
  saveRDS(hop_phylo, 
        "Hopfner_Species.rds")
}

```


### Keshavarzian et al
```{r Keshavarzian}
# Load Species Count table
kes<-read.delim("Species.txt", header = T, stringsAsFactors = F)
row.names(kes)<-kes$Species
kes<-kes[,-1]

# Load Lotus mapping file
map<-read.delim("Keshavarz_et_al_mapping.txt", header = T, stringsAsFactors = F)
names(map)[1]<-"SampleID"

# Check that the samples order is the same in count table and mapping file
kes<-kes[,order(colnames(kes))]
map<-map[order(map$SampleID),]
all(colnames(kes) == map$SampleID)

# Create phyloseq object
rownames(map)<-map$SampleID
map<-map[,-1]
kes_phylo <- phyloseq(otu_table(as.matrix(kes), taxa_are_rows = T), 
                        sample_data(map))

# Remove taxa that have < 5 counts in < 2.5% taxa
wh0 <- genefilter_sample(kes_phylo, filterfun_sample(function(x) x > 5), A=0.025*nsamples(kes_phylo))
kes_phylo <- prune_taxa(wh0, kes_phylo)

# Check whether there are taxa in >= 90% samples
df<-as.data.frame(otu_table(kes_phylo))
df[df>=1]<-1
if(length(rowSums(df)[rowSums(df) >= 0.9*ncol(df)]) >1){
  print("There are taxa in >90% of samples")
  length(rowSums(df)[rowSums(df) >= 0.9*ncol(df)])
}else{stop("Something is wrong here, there are not taxa present in > 90% samples")}

# Look at hte number of PD and HC samples

sum(get_variable(kes_phylo)$Description == "Parkinson") 
sum(get_variable(kes_phylo)$Description == "Control") 

# Remove samples with < 4500 reads
reads<-read.delim("demulti.acceptsPerSample.log", header = T, stringsAsFactors = F)

if(nrow(subset(reads, Read1Accepted < 4500)) == 0){
  saveRDS(kes_phylo, 
        "Keshavarz_Species.rds")
}else{
  keep<-subset(reads, Read1Accepted > 4500)$SampleID
  kes_phylo <-prune_samples(keep, kes_phylo)
  if(any(taxa_sums(kes_phylo) == 0)){
    kes_phylo <- prune_taxa(taxa_sums(kes_phylo) > 0, kes_phylo)
    }
  saveRDS(kes_phylo, 
        "Keshavarz_Species.rds")
}
```

### Create a combined dataset
```{r, Combined dataset}
# Create Scheperjans and Aho separate datasets
sch_base<-subset_samples(sa_phylo, row.names(sample_data(sa_phylo)) %in% sample_names(sa_phylo)[grepl("baseline", sample_names(sa_phylo))])
aho_foll<-subset_samples(sa_phylo, row.names(sample_data(sa_phylo)) %in% sample_names(sa_phylo)[grepl("followup", sample_names(sa_phylo))])

# Look at the number of PD and HC samples in Aho et al baseline and followup datsets

sum(get_variable(sch_base)$Parkinson == "Parkinson") 
sum(get_variable(sch_base)$Parkinson == "Control") 

sum(get_variable(aho_foll)$Parkinson == "Parkinson") 
sum(get_variable(aho_foll)$Parkinson == "Control") 


# Check num are 1211
nsamples(pietr_phylo) + nsamples(qian_phylo) + nsamples(hb_phylo) + nsamples(heinz_phylo) + 
  nsamples(petr_phylo) + nsamples(weis_phylo) + nsamples(sch_base) + nsamples(aho_foll) + nsamples(hop_phylo) +
  nsamples(kes_phylo)

# Standardize the formatting of the mapping files
map_schaho<-sample_data(sa_phylo)[,"Parkinson"]
map_sch<-map_schaho[grepl("baseline", rownames(map_schaho)),]
map_sch$study<-rep("Scheperjans", nrow(map_sch))
names(map_sch)<-c("status", "study")

map_aho<-map_schaho[grepl("followup", rownames(map_schaho)),]
map_aho$study<-rep("Aho", nrow(map_aho))
names(map_aho)<-c("status", "study")

map_pietr<-sample_data(pietr_phylo)
map_pietr_1<-map_pietr[, "host_phenotype"]
map_pietr_1$Study<-rep("Pietrucci", nrow(map_pietr_1))
names(map_pietr_1)<-c("status", "study")

map_qian<-sample_data(qian_phylo)
map_qian_1<-map_qian[, "status"]
map_qian_1$Study<-rep("Qian", nrow(map_qian_1))
names(map_qian_1)<-c("status", "study")

map_hillb<-sample_data(hb_phylo)
map_hillb_1<-map_hillb[, "Parkinson"]
map_hillb_1$Study<-rep("Hill-Burns", nrow(map_hillb_1))
names(map_hillb_1)<-c("status", "study")

map_heinz<-sample_data(heinz_phylo)
map_heinz_1<-map_heinz[, "Status"]
map_heinz_1$Study<-rep("Heintz-Buschart", nrow(map_heinz_1))
names(map_heinz_1)<-c("status", "study")

map_petrov<-sample_data(petr_phylo)
map_petrov_1<-map_petrov[, "diagnosis"]
map_petrov_1$Study<-rep("Petrov", nrow(map_petrov_1))
names(map_petrov_1)<-c("status", "study")

map_weis<-sample_data(weis_phylo)
x<-rownames(map_weis)
for(i in 1:length(x)){
  if(grepl("mK", x[i])){
    x[i]<-"Control"
  }else{
      x[i]<-"Parkinson"
  }
}

map_weis_1<-data.frame(status = x, study = rep("Weis", nrow(map_weis)))
rownames(map_weis_1)<-sample_names(weis_phylo)
names(map_weis_1)<-c("status", "study")

# Calculate number of samples for Weis

sum(map_weis_1$status == "Parkinson")
sum(map_weis_1$status == "Control")

map_hop<-sample_data(hop_phylo)
map_hop_1<-map_hop[, "Status"]
map_hop_1$Study<-rep("Hopfner", nrow(map_hop_1))
names(map_hop_1)<-c("status", "study")

map_kev<-sample_data(kes_phylo)
map_kev_1<-map_kev[, "Description"]
map_kev_1$Study<-rep("Keshavarzian", nrow(map_kev_1))
names(map_kev_1)<-c("status", "study")

map_combined<-rbind.data.frame(map_heinz_1, map_hillb_1, map_petrov_1, map_weis_1, map_pietr_1, 
                               map_qian_1, map_sch, map_aho, map_hop_1, map_kev_1)

# Check whether there are any duplicate samples in the map
if(nrow(map_combined[duplicated(rownames(map_combined))]) == 0){
  stop("Formatting of the mapping files was not done correctly")
}else{print("Formatting of the mapping files was done correctly")}

# Marging count table to creare a combined (pooled) dataset
d_heinz<-otu_table(heinz_phylo)
d_pietr<-otu_table(pietr_phylo)
d_petrov<-otu_table(petr_phylo)
d_qian<-otu_table(qian_phylo)
d_sch<-otu_table(sch_base)
d_aho<-otu_table(aho_foll)
d_kev<-otu_table(kes_phylo)
d_weis<-otu_table(weis_phylo)
d_hop<-otu_table(hop_phylo)
d_hill<-otu_table(hb_phylo)

# Check that the number of samples is correct (1211) and there are no duplicates
sum(nsamples(d_hill), nsamples(d_heinz), nsamples(d_pietr), nsamples(d_petrov),
    nsamples(d_qian), nsamples(d_sch), nsamples(d_aho), nsamples(d_kev), nsamples(d_weis),
    nsamples(d_hop))

nsam<-(c(sample_names(d_hill), sample_names(d_heinz), sample_names(d_pietr), sample_names(d_petrov),
         sample_names(d_qian), sample_names(d_sch), sample_names(d_aho), sample_names(d_kev), sample_names(d_weis),
         sample_names(d_hop)))
if(sum(duplicated(nsam)) != 0){
  stop("Formatting of the mapping files was not done correctly")
}else{print("Formatting of the mapping files was done correctly")}

# Merging the count tables
hh<-merge(d_hill, d_heinz, by = 0, all = T)
row.names(hh)<-hh$Row.names
hhp<-merge(hh[,-1], d_pietr, by = 0, all = T)
row.names(hhp)<-hhp$Row.names
p<-merge(hhp[,-1], d_petrov, by = 0, all = T)
row.names(p)<-p$Row.names
q<-merge(p[,-1], d_qian, by = 0, all = T)
row.names(q)<-q$Row.names
s<-merge(q[,-1], d_sch, by = 0, all = T)
row.names(s)<-s$Row.names
sk<-merge(s[,-1], d_kev, by = 0, all = T)
row.names(sk)<-sk$Row.names
skw<-merge(sk[,-1], d_weis, by = 0, all = T)
row.names(skw)<-skw$Row.names
skwa<-merge(skw[,-1], d_aho, by = 0, all = T)
row.names(skwa)<-skwa$Row.names
d_comb<-merge(skwa[,-1], d_hop, by = 0, all = T)
row.names(d_comb)<-d_comb$Row.names
d_comb<-d_comb[,-1] 
# Check that the number of samples is correct (1211)
ncol(d_comb)

# Change NA in 0, format otu_tab and sample_data in the same order of samples, create phyloseq object
d_comb[is.na(d_comb)]<-0

d_comb<-d_comb[, order(names(d_comb))]
map_combined<-map_combined[order(rownames(map_combined)),]
all(names(d_comb) == rownames(map_combined))

map_combined$study<-as.factor(map_combined$study)
d_phylo_com_count<-phyloseq(otu_table(d_comb, taxa_are_rows = T), sample_data(map_combined))

# Create a list of phyloseq objects
l_phylo<-list(pietr_phylo, qian_phylo, 
              hb_phylo, heinz_phylo, 
              petr_phylo, weis_phylo, 
              sch_base, aho_foll, 
              hop_phylo, kes_phylo, 
              d_phylo_com_count)
names(l_phylo)<-c("Pietrucci", "Qian", "Hill-Burns", "Heintz-Buschart", "Petrov", "Weis", "Scheperjans", 
                  "Aho", "Hopfner","Keshavarzian", "Combined")

# Create a list of maps and update the current mapping files of the phyloseq object in the list
l_maps<-list(map_pietr_1, map_qian_1, map_hillb_1, map_heinz_1, map_petrov_1, map_weis_1, map_sch, map_aho, 
             map_hop_1, map_kev_1, map_combined)
names(l_maps)<-names(l_phylo)

for(i in 1:length(l_maps)){
  if(all(rownames(sample_data(l_phylo[[i]])) == rownames(l_maps[[i]]))){
    sample_data(l_phylo[[i]])<-sample_data(l_maps[[i]])
  }else{stop("Order of samples is different! Re-order the phyloseq samples and the mapping file")}
}

saveRDS(l_maps,
        "Formatted_maps.rds")
saveRDS(l_phylo, 
        "Phyloseq_list_all_studies.rds")
```
