---
title: "Alpha_diversity_species"
author: "Stefano Romano"
output: html_document
editor_options: 
  chunk_output_type: console
---
# Before you begin:

These scripts were tailored for the analyses we conducted. They are not meant to be used in other contexts and are reported here as a detailed extension of the method section of our article. Please, use them at your own risk.

If you use any of the material or approaches in these scripts, please, cite:
Romano et al, 2020, *Meta-analysis of the Parkinson’s disease gut microbiome suggests alterations linked to intestinal inflammation*

Thank you!


# Load the needed packages
```{r, message=F}
list.of.packages <- c("ggplot2", "reshape2", "dplyr", "devtools", "ggpubr", "data.table", "genodds", "meta", "grid")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

if (!requireNamespace("BiocManager", quietly = TRUE)){
  install.packages("BiocManager")}


list.of.bioc <- c("phyloseq", "microbiome")
new.packages <- list.of.bioc[!(list.of.bioc %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)


library(phyloseq)
library(microbiome)
library(ggpubr)
library(ggplot2)
library(reshape2)
library(dplyr)
library(meta)
library(data.table)
library(genodds)
library(grid)


```

# Alpha diversity analyses
We will calculate various alpha diversity indices and test whether their differences between PD and controls are statistically significant. We will use for testing purposes Agresti's generalized odds ratios and a mixed-effect meta-analysis approach to pool the results.

A full path to the R objects and files should be provided.

## Load the list pf phyloseq objects
```{r setup, message=FALSE}
l_phylo<-readRDS("Phyloseq_list_all_studies.rds")
```

## Remove Eukaryotes and Choroplasts and rarefy.
No sequences calssified as Mitochondria were found in the datasets

```{r, message=F}
# Remove Eukaryotes
l_phylo<-lapply(l_phylo, function(x) prune_taxa(taxa_names(x)[!(grepl("Eukaryota", taxa_names(x), ignore.case = TRUE))], x))

# Check that there are not Euk left
if(length(unlist(lapply(l_phylo, function(x) taxa_names(x)[grep("Eukary*", taxa_names(x))]))) == 0){
  print("All Eukaryotes removed")
}else{stop("There are still Eukaryotes sequences in the files")}

# Check Chloroplasts and Mitochondria
for(i in 1:length(l_phylo)){
  if(sum(grep("Chloroplast", taxa_names(l_phylo[[i]]))) > 0){
    l_phylo[[i]]<-prune_taxa(taxa_names(l_phylo[[i]])[!(grepl("Chloropl*", taxa_names(l_phylo[[i]]), ignore.case = TRUE))], l_phylo[[i]])
  } else {}
}

if(length(unlist(lapply(l_phylo, function(x) taxa_names(x)[grep("Chloropl*", taxa_names(x))]))) == 0){
  print("All Chloroplast removed")
}else{stop("There are still Chloroplast sequences in the files")}


if(length(unlist(lapply(l_phylo, function(x) taxa_names(x)[grep("Mitoch*", taxa_names(x))]))) == 0){
  print("All mitochondria removed")
}else{stop("There are still mitochondria sequences in the files")}
# In all dataets there were no mitochondria sequences

# Remove eventual samples with 0 sequence
l_phylo<-lapply(l_phylo, function(x) prune_samples(sample_sums(x) > 0, x))


# Rarefy to 5000, no replacements, OTUs no longer observed in any sample are removed
# Samples with fewer reads than samples size (5000) will be removed

l_rar<-lapply(l_phylo, function(x) rarefy_even_depth(x, 
                                                    sample.size = 5000, 
                                                    replace = FALSE, 
                                                    trimOTUs = TRUE, 
                                                    rngseed = 56987))
# Save rarefied data for the other analyses
saveRDS(l_rar, "raref_list_all_studies.rds")
```

## Calculate Firmicutes Bacteroidota ratio (F/B) ratio.

The log2 ratio is calcualte using the relative abudance of each taxa

```{r, message=F}
# Remove the combined datsets
l_rar<-l_rar[-11] # Remove the combine dataset

# Create a file with proportions
l_prop<-lapply(l_rar, function(x) transform_sample_counts(x, function(x){x / sum(x)}))

# Calcualte F/B ratio from Species table
firm<-vector(mode = "list")
bact<-vector(mode = "list")
fbratio<-vector(mode  ="list", length = length(l_prop))
for(i in 1:length(l_prop)){
  firm[[i]]<-as.data.frame(otu_table(l_prop[[i]])[grep("Firmicutes", rownames(otu_table(l_prop[[i]]))),])
  bact[[i]]<-as.data.frame(otu_table(l_prop[[i]])[grep("Bacteroidota", rownames(otu_table(l_prop[[i]]))),])
  if(all(names(firm[[i]]) == names(bact[[i]])) == TRUE){
    print("Firma and Bact. table have the same order of samples")
  }
  else{
    (stop("Order of samples is different in Firm and Bact table"))
  }
  for(s in 1:ncol(firm[[i]])){
    fbratio[[i]]$sample[s]<-names(firm[[i]])[s]
    fbratio[[i]]$log2_fb_ratio[s]<-log(sum(firm[[i]][,s])/sum(bact[[i]][,s]), base = 2)
  }
  fbratio[[i]]<-data.frame(sample = fbratio[[i]]$sample, log2_fb_ratio = fbratio[[i]]$log2_fb_ratio)
}
```

## Calculate Prevotella Bacteroides ratio (P/B) ratio.
The log2 ratio is calculated using the relative abudance of each taxa

```{r, message=F}
# Calcualte F/B ratio from Species table
prev<-vector(mode = "list")
bact<-vector(mode = "list")
pbratio<-vector(mode  ="list", length = length(l_prop))
for(i in 1:length(l_prop)){
  prev[[i]]<-as.data.frame(otu_table(l_prop[[i]])[grep("Prevotella", rownames(otu_table(l_prop[[i]]))),])
  bact[[i]]<-as.data.frame(otu_table(l_prop[[i]])[grep("Bacteroides", rownames(otu_table(l_prop[[i]]))),])
  if(all(names(prev[[i]]) == names(bact[[i]])) == TRUE){
    print("Prev and Bact. table have the same order of sample")
  }
  else{
    (stop("Order of samples is different in Prevo and Bact table"))
  }
  for(s in 1:ncol(prev[[i]])){
    pbratio[[i]]$sample[s]<-names(prev[[i]])[s]
    pbratio[[i]]$log2_pb_ratio[s]<-log(sum(prev[[i]][,s])/sum(bact[[i]][,s]), base = 2)
  }
  pbratio[[i]]<-data.frame(sample = pbratio[[i]]$sample, log2_pb_ratio = pbratio[[i]]$log2_pb_ratio)
}
```

## Calculate the rest of the alpha indeces

```{r, message=FALSE}
ind<-c("observed",  "chao1", "diversity_fisher", 
       "evenness_simpson", "evenness_bulla", "dominance_core_abundance",
       "dominance_simpson", "rarity_low_abundance", "rarity_rare_abundance")

l_rich<-lapply(l_rar, function(x) microbiome::alpha(x, ind))
names(l_rich)<-names(l_rar)

# Add the ACE from phyloseq
for(i in 1:length(l_rich)){
  ps<-estimate_richness(l_rar[[i]], measures = c("ACE"))
  l_rich[[i]]$ACE<-ps$ACE
}


l_maps<-lapply(l_rar, function(x) meta(x))
names(l_maps)<-names(l_rar)  
# Combine the metadata and the alpha diversity indices
rich_map<-l_maps
for(i in 1: length(l_maps)){
    l_rich[[i]]<-l_rich[[i]][order(rownames(l_rich[[i]])),]
    l_maps[[i]]<-l_maps[[i]][order(rownames(l_maps[[i]])),]
    fbratio[[i]][order(fbratio[[i]]$sample),]
    if(all(rownames(l_rich[[i]]) == fbratio[[i]]$sample)){
        print("Order of samples in the two df is the same, fb_ratio and other indices ca be merged")
        l_rich[[i]]$log2_fb_ratio<-fbratio[[i]]$log2_fb_ratio
      }else{stop("There is a problem, samples in files are not in the same order")
    }
     if(all(rownames(l_rich[[i]]) == pbratio[[i]]$sample)){
        print("Order of samples in the two df is the same, pb_ratio and other indices ca be merged")
        l_rich[[i]]$log2_pb_ratio<-pbratio[[i]]$log2_pb_ratio
      }else{stop("There is a problem, samples in files are not in the same order")
    }
    if(all(rownames(l_rich[[i]]) == rownames(l_maps[[i]]))){
      rich_map[[i]]<-cbind(l_rich[[i]], l_maps[[i]])
      }else{stop("Richness and meta data have samples in diff order. Adjust this before proceeding")}
    
}
```

# Agresti's generalized odds ratios 
```{R}
# Create a df and format it for the following analyses
tab1<-do.call(rbind.data.frame, rich_map) 
tab1$samples<-rownames(tab1)
tab1df <- data.table(tab1)
tab1$status <- as.factor(tab1$status)
tab1$study <- as.factor(tab1$study)

# Meta analysis using Agresti's log-generalized odds ratio.
meta_alphas2 <- lapply(1:12, function(a){
  print(a)
  results <- by(data=data.frame(tab1), INDICES = tab1$study, FUN = function(d) {
    lo <- genodds(d[,a], d$status)
    c(lo$pooled_lnodds, lo$pooled_SElnodds)
    }
    )
  results <- data.frame(do.call(rbind,results))
  names(results) <- c("estimate", "se")
  results$study <- levels(tab1$study)
  # For each measure, return a list that is the meta-analysis results [[1]] and the
  # table of individual results in [[2]]
  list(metagen(TE=estimate, seTE=se,data=results,method.tau = "ML",studlab=study, comb.fixed = F, comb.random = T), 
       results)
})

names(meta_alphas2)<-names(tab1)[1:12]

## Now gather all the results together
random.meta <- do.call(rbind,lapply(meta_alphas2, function(m) as.data.frame(m[[1]][c("TE.random","seTE.random", "lower.random", "upper.random", "pval.random")])))
random.meta$statistic <- names(tab1)[1:12]
random.meta$statistic <- factor(random.meta$statistic, levels=random.meta$statistic)


# Do two series of graphs, one for diversity indices and one for f/b and p/b ratios
# Format the dfs
ind<-c("ACE", ind)
df<-subset(random.meta, statistic %in% ind)

df$statistic<-factor(df$statistic, levels = c("observed", "chao1","ACE", 
                                              "diversity_fisher", "evenness_simpson",
                                              "evenness_bulla", "dominance_core_abundance",
                                              "dominance_simpson", "rarity_low_abundance",
                                              "rarity_rare_abundance"),
                     labels = c("Observed", "Chao1","ACE", 
                                "Fisher alpha", "Evenness Simpson",
                                              "Evenness Bulla", "Dominance core abundance",
                                              "Dominance Simpson", "Rarity low abundance",
                                              "Rarity rare abundance"))
p<-ggplot(df, aes(y=TE.random, 
                        ymax=upper.random, 
                        ymin=lower.random,
                        x=statistic)) + 
  geom_errorbar(width=0.2) + 
  geom_point(aes(fill = seTE.random), shape = 21, size =4.5) + 
  scale_fill_viridis_c(breaks = c(0.08, 0.09,0.1, 0.11)) +
  theme_bw() + 
  coord_flip() + 
  geom_hline(yintercept=0, linetype="dashed", color="gray50") + 
  labs(y="Log-Generalised Odds Ratio (95% CI)", 
      x="Diversity Indices",
      fill = "Standard\nerror") +
  theme(axis.title.x = element_text(size =16, face = "plain", vjust = 1),
        axis.title.y = element_text(size =16, face = "plain"),
        axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size =16), 
        legend.title = element_text(size =14, face = "bold"),
        legend.text = element_text(size = 14),
        plot.margin = margin(35,1,1,1)) 

library(grid)
library(gridExtra)

# Create a text and title to add
grob <- grobTree(textGrob("Control", x=0.01,  y=0.02, hjust=0,
  gp=gpar(col="black", fontsize=14, fontface="italic")))
grob2 <- grobTree(textGrob("PD", x=0.92,  y=0.02, hjust=0,
  gp=gpar(col="black", fontsize=14, fontface="italic")))

(p2<-p + annotation_custom(grob) + 
  annotation_custom(grob2))

p1<-ggdraw(p2) + draw_text("Changes in diversity indices between PD and controls", 
                      x = 0.27, y = 0.98, hjust = 0, vjust = 1,
                      size = 18) 

# Save figure
ggsave2("Alpha.svg", 
        p1, 
        device = "svg", 
        width = 8.3, 
        height = 7, 
        dpi = 300)

# Make a forest plot for each index 
# Format the results in a df
metan<-lapply(meta_alphas2, function(x) x[[1]])
meta2df<-function(meta){
  df<-data.frame(study = vector(), estimate = vector(), se_est = vector(),
                 lower = vector(), upper = vector(), w.random = vector(),
                 index = vector(), het = vector())
    
  for(i in 1:length(meta)){
    lab<-meta[[i]]$studlab
    index<-rep(names(meta)[i], length(meta[[i]]$studlab))
    te<-meta[[i]]$TE
    se<-meta[[i]]$seTE
    low<-meta[[i]]$lower
    up<-meta[[i]]$upper
    weight<-meta[[i]]$w.random
    d<-data.frame(study = lab, estimate = te, se_est = se,
                 lower = low, upper = up, w.random = weight,
                 index = index, het = rep(NA, length(index)))
    d$study<-as.character(d$study)
    d[(nrow(d)+1),1]<-"Random effect model"
    d[nrow(d),2:6]<-c(meta[[i]]$TE.random,meta[[i]]$seTE.random,meta[[i]]$lower.random,meta[[i]]$upper.random,100)
    d[nrow(d),7]<-names(meta)[i]
    I2<-round(meta[[i]]$I2, 3)*100
    tu<-round(metan[[i]]$tau2,2)
    pval<-round(metan[[i]]$pval.Q, 2)
    d[nrow(d),8]<-paste0("Het: I","\u00b2", "=", I2, ", tau","\u00b2","=", tu, ", p-val=", pval)
    d$study<-factor(d$study, levels = unique(d$study))
    d$study<-relevel(d$study, "Random effect model")
    df<-rbind.data.frame(df,d)
  }
  return(df)
}

# Plot forest plot in ggplot
df<-meta2df(metan)
df$w.random[df$w.random ==100] <-NA
a<-ggplot(df[df$index%in% ind,], aes(y=estimate, 
                        ymax=upper, 
                        ymin=lower,
                        x=study)) + 
  geom_errorbar(width=0.2) + 
  geom_point(aes(fill = se_est, size = w.random), shape = 21) + 
  scale_fill_viridis_c() +
  theme_bw() + 
  coord_flip() + 
  facet_wrap(~index, ncol = 5) +
  geom_hline(yintercept=0, linetype="dashed", color="gray50") + 
  labs(fill = "SE",
       size = "Weights in random effect model [%]",
       x = "Study",
       y = "Log-Generalised Odds Ratio (95% CI)",
       title = "Changes in diversity indices between PD and Controls") +
  theme(axis.title.x = element_text(size =16, face = "plain", vjust = 1),
        axis.title.y = element_text(size =16, face = "plain"),
        axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size =16), 
        legend.title = element_text(size =14, face = "bold"),
        legend.text = element_text(size = 14),
        strip.text = element_text(size = 14),
        title = element_text(size = 18, face = "bold"),
        legend.position = "bottom")


# Add text
tag_facet2 <- function(p, open = "(", close = ")", tag_pool = letters, x = -Inf, y = Inf, 
    hjust = -0.5, vjust = 1.5, fontface = 2, family = "", ...) {

    gb <- ggplot_build(p)
    lay <- gb$layout$layout
    tags <- cbind(lay, label = paste0(open, tag_pool[lay$PANEL], close), x = x, y = y)
    p + geom_text(data = tags, aes_string(x = "x", y = "y", label = "label"), ..., hjust = hjust, 
        vjust = vjust, fontface = fontface, family = family, inherit.aes = FALSE)
}

df2<-subset(df,index %in% ind )
my_tag <- df2$het[!(is.na(df2$het))]
at<-tag_facet2(a, 
          x = -Inf, y = -Inf, 
          vjust = -0.2, hjust = -0.01,
          open = "", close = "",
          fontface = 1,
          size = 4,
          family = "serif",
          tag_pool = my_tag)

ggsave(at, 
       filename = "Alpha_forest.tiff", 
       device = "tiff", 
       width = 16, 
       height = 11, 
       dpi = 300)



# Do graph for f/b and p/b

(a<-ggplot(df[df$index%in% c("log2_fb_ratio", "log2_pb_ratio"),], aes(y=estimate, 
                        ymax=upper, 
                        ymin=lower,
                        x=study)) + 
  geom_errorbar(width=0.2) + 
  geom_point(aes(fill = se_est, size = w.random), shape = 21) + 
  scale_fill_viridis_c() +
  theme_bw() + 
  coord_flip() + 
  facet_wrap(~index, ncol = 5) +
  geom_hline(yintercept=0, linetype="dashed", color="gray50") + 
  labs(fill = "SE",
       size = "Weights in random effect model [%]",
       x = "Study",
       y = "Log-Generalised Odds Ratio (95% CI)",
       title = "Changes in diversity indices between PD and Controls") +
  theme(title = element_text(size = 15, face = "bold"),
        axis.title = element_text(size =14, face = "plain"),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size =14), 
        legend.title = element_text(size =12, face = "bold"),
        legend.text = element_text(size = 12),
        legend.position = "bottom",
        legend.direction = "horizontal",
        strip.text = element_text(size = 14))) 

# Add text
df2<-subset(df,index %in% c("log2_fb_ratio", "log2_pb_ratio") )
my_tag <-  unique(df2$het)[-1]
at<-tag_facet2(a, 
          x = -Inf, y = -Inf, 
          vjust = -0.2, hjust = -0.01,
          open = "", close = "",
          fontface = 1,
          size = 4,
          family = "serif",
          tag_pool = my_tag)

# Save  
ggsave(at, 
       filename = "Pb_fb.tiff", 
       device = "tiff", 
       width = 9, 
       height = 5, 
       dpi = 300)

```

## Save workspace
```{r, message=FALSE}
save.image("Alpha_species.RData")
```
