---
title: "VST_Beta_DESeq2_DA"
author: "Stefano Romano"
output: html_document
editor_options: 
  chunk_output_type: console
---
# Before you begin:

These scripts were tailored for the analyses we conducted. They are not meant to be used in other contexts and are reported here as a detailed extension of the method section of our article. Please, use them at your own risk.

If you use any of the material or approaches in these scripts, please, cite:
Romano et al, 2020, *Meta-analysis of the Parkinson’s disease gut microbiome suggests alterations linked to intestinal inflammation*

Thank you!

# Load the needed packages
```{r, message=F}
list.of.packages <- c("ggplot2", "dplyr", "reshape2", "devtools", "ggpubr", "vegan")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

if (!requireNamespace("BiocManager", quietly = TRUE)){
  install.packages("BiocManager")}


list.of.bioc <- c("DESeq2", "phyloseq",  "grDevices", "microbiome")
new.packages <- list.of.bioc[!(list.of.bioc %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)


library(phyloseq)
library(ggpubr)
library(DESeq2)
library(ggplot2)
library(reshape2)
library(dplyr)
library(microbiome) 
library(grDevices)
library(devtools)
library(vegan)
source('GMPR.R') # Version: 0.1. Date: 2017/02/07
```

# Overview of the analyses
We will verify the effect of disease status and study on the microbial community structure and perform a differential abundance analysis using DESeq2.
Please provide the full path to the files and R objects that need to be loaded.

## Load the list pf phyloseq objects
```{r setup, message=FALSE}
l_phylo<-readRDS("Phyloseq_list_all_studies.rds")
```

## Remove Eukaryotes and Chloroplast. Create DeSeq objects
```{r, message=F}
# Remove Eukaryotes
l_phylo<-lapply(l_phylo, function(x) prune_taxa(taxa_names(x)[!(grepl("Eukaryota", taxa_names(x), ignore.case = TRUE))], x))

# Check that there are not Euk left
if(length(unlist(lapply(l_phylo, function(x) taxa_names(x)[grep("Eukary*", taxa_names(x))]))) == 0){
  print("All Eukaryotes removed")
}else{stop("There are still Eukaryotes sequences in the files")}

# Check Chloroplasts
for(i in 1:length(l_phylo)){
  if(sum(grep("Chloroplast", taxa_names(l_phylo[[i]]))) > 0){
    l_phylo[[i]]<-prune_taxa(taxa_names(l_phylo[[i]])[!(grepl("Chloropl*", taxa_names(l_phylo[[i]]), ignore.case = TRUE))], l_phylo[[i]])
  } else {}
}

if(length(unlist(lapply(l_phylo, function(x) taxa_names(x)[grep("Chloropl*", taxa_names(x))]))) == 0){
  print("All Chloroplast removed")
}else{stop("There are still Chloroplast sequences in the files")}

# Remove samples with 0 counts if any 
l_phylo<-lapply(l_phylo, function(x) prune_samples(sample_sums(x) > 0, x))

# transform variable in factor
for(i in 1:length(l_phylo)){
  x<-meta(l_phylo[[i]])
  x$status<-as.factor(x$statu)
  x$study<-as.factor(x$study)
  sample_data(l_phylo[[i]])<-sample_data(x)
}

# Make DESeq objects
l_deseq<-lapply(l_phylo, function(x) phyloseq_to_deseq2(x, ~ status)) # The warning here is meaningless because we are going to create a new ds object below for the combined dataset
l_deseq$Combined<-phyloseq_to_deseq2(l_phylo[["Combined"]], design = ~ study + status)

```

## Calcualte Size facor and perform DeSeq analyses

```{r, message=F}
for(i in 1:length(l_deseq)){
  gmpr.size.factor <- GMPR(as.matrix(as.data.frame(otu_table(l_phylo[[i]]))))
  sizeFactors(l_deseq[[i]])<-gmpr.size.factor
}

l_dseq_ds<-lapply(l_deseq, function(x) DESeq(x))
l_deseq_res<-lapply(l_dseq_ds, function(x) results(x, independentFiltering  = TRUE))
nrow(l_deseq_res[["Combined"]])
l_deseq_sign<-lapply(l_deseq_res, function(x) subset(x, padj <= 0.05))
nrow(l_deseq_sign[["Combined"]])

# Create a list with the name of the sign taxa 
l_sign_deseq_taxa<-lapply(l_deseq_sign, function(x) rownames(x))
saveRDS(l_sign_deseq_taxa, "DeSeq2_GMPR_list_Species_DA.rds")

```


## Perform beta diversity analyses and clustering

```{r, message=FALSE}
set.seed(987)
# Since there are too many 0 then I use the GMPR transformations (re-called from above dseq object)
l_vst<-lapply(l_deseq, function(x) varianceStabilizingTransformation(x, blind = FALSE))

# Add the tranformed data to te phyloseq objects
for(i in 1:length(l_vst)){
  otu_table(l_phylo[[i]])<-otu_table(assay(l_vst[[i]]), taxa_are_rows = T)
}

# Calculate indices 
l_bray<-lapply(l_phylo, function(x) phyloseq::distance(x, method= "bray", na.rm = F))
l_jsd<-lapply(l_phylo, function(x) phyloseq::distance(x, method= "jsd", na.rm=F))
l_euc<-lapply(l_phylo, function(x) phyloseq::distance(x, method= "euclidean", na.rm=F))

# Calculate adonis for the 3 indices
l_adonis_b<-l_bray
names(l_adonis_b)<-names(l_bray)
for(i in 1:(length(l_bray)-11)){
  l_adonis_b[[i]]<-adonis2(l_bray[[i]] ~ status, data = meta(l_phylo[[i]]),add = "lingoes")
}
  l_adonis_b[[11]]<-adonis2(l_bray[[11]] ~ status, data = meta(l_phylo[[11]]))

l_adonis_j<-l_jsd
names(l_adonis_j)<-names(l_jsd)
for(i in 1:(length(l_bray)-11)){
  l_adonis_j[[i]]<-adonis2(l_jsd[[i]] ~ status, data = meta(l_phylo[[i]]),add = "lingoes")
}
  l_adonis_j[[11]]<-adonis2(l_jsd[[11]] ~ status, data = meta(l_phylo[[11]]))

l_adonis_e<-l_euc
names(l_adonis_e)<-names(l_euc)
for(i in 1:length(l_euc)){
  l_adonis_e[[i]]<-adonis2(l_euc[[i]] ~ status, data = meta(l_phylo[[i]]))
}

# Create a df with the adonis values 
adonis_df_b<-data.frame(matrix(nrow = length(l_adonis_b), ncol = 4))
for(i in 1:length(l_bray)){
  adonis_df_b[i,]<- c(names(l_adonis_b)[i], l_adonis_b[[i]]$R2[1], l_adonis_b[[i]]$`Pr(>F)`[1], "Bray_VST")
}
adonis_df_jsd<-data.frame(matrix(nrow = length(l_adonis_b), ncol = 4))
for(i in 1:length(l_bray)){
  adonis_df_jsd[i,]<- c(names(l_adonis_j)[i], l_adonis_j[[i]]$R2[1], l_adonis_j[[i]]$`Pr(>F)`[1], "JSD_VST")
}
adonis_df_euc<-data.frame(matrix(nrow = length(l_adonis_e), ncol = 4))
for(i in 1:length(l_bray)){
  adonis_df_euc[i,]<- c(names(l_adonis_e)[i], l_adonis_e[[i]]$R2[1], l_adonis_e[[i]]$`Pr(>F)`[1], "Euclidean_VST")
}

# Combine and save results for plotting
adonis_df<-rbind.data.frame(adonis_df_b, adonis_df_jsd, adonis_df_euc)
names(adonis_df)<-c("Study", "R2", "pvalue", "Index")


for(i in 1:nrow(adonis_df)){
  if(adonis_df$pvalue[i] < 0.1 & adonis_df$pvalue[i] > 0.05){
    adonis_df$Significant[i]<-"Significant [0.05<p<0.1]"
  } else if (adonis_df$pvalue[i] < 0.05){
    adonis_df$Significant[i]<-"Significant [p<0.05]"
  } else {adonis_df$Significant[i]<-"Not_significant"
  }
}

adonis_df$Significant<-as.factor(adonis_df$Significant)
adonis_df$R2<-round(as.numeric(adonis_df$R2)*100, digits = 1)

saveRDS(adonis_df, "Adonis_species_VST.rds")

# Check the marginal effect of study and disease status on the 3 different distances
adonis_com<-adonis2(l_jsd$Combined ~ study + status, data = meta(l_phylo$Combined), by = "margin")
adonis_com_b<-adonis2(l_bray$Combined ~ study + status, data = meta(l_phylo$Combined), by = "margin")
adonis_com_e<-adonis2(l_euc$Combined ~ study + status, data = meta(l_phylo$Combined), by = "margin")


# dbRDA (capscale) using Euclidean
l_dbRDA_spe<-vector(mode = "list", length = 3)
l_dbRDA_spe[[1]]<-ordinate(l_phylo[["Combined"]], method = "CAP", distance = l_euc[["Combined"]], formula = l_phylo[["Combined"]] ~ 1 )
l_dbRDA_spe[[2]]<-ordinate(l_phylo[["Combined"]], method = "CAP", distance = l_euc[["Combined"]], formula = l_phylo[["Combined"]] ~ 1  + Condition(study))
l_dbRDA_spe[[3]]<-ordinate(l_phylo[["Combined"]], method = "CAP", distance = l_euc[["Combined"]], formula = l_phylo[["Combined"]] ~  status + Condition(study))

names(l_dbRDA_spe)<-c("Constrained for Study and Status", "Blocked", "Blocked_constrained")
l_dbRDA_spe_plots<-lapply(l_dbRDA_spe, function(x) plot_ordination(l_phylo[["Combined"]], x, 
                "samples", color="status", shape = "study") +
                theme_bw() +
                scale_shape_manual(values=c(2,5,6,11,17,0,19,7,8,4)) +
                scale_colour_manual(values=c("#999999", "#E69F00")) +
                geom_point(size = 2.5) +
                labs(shape = "Study", color = "Status") +
                theme(legend.title = element_text(size = 12, face = "bold"),
                      legend.text = element_text(size = 12)))
              
                
g<-ggarrange(plotlist = l_dbRDA_spe_plots, common.legend = T, font.label = list(size = 10), label.x = 0.15, label.y = 0.975, ncol = 3,
             labels = c("a: Unconstrained", "b: Blocked by study", "c: Blocked by study constrained by condition"),
             hjust = c(0,0,0),legend = "bottom") 

g_Species<-annotate_figure(g, top = text_grob("Species", size = 12, face = "bold"))

# Save RDS
saveRDS(g_Species, "dbRDA_species_VST.rds")
```

## Save workspace
```{r, message=FALSE}
save.image("04_Species_DESEQ_GMPR_workspace.RData")

```
