---
title: "DA_ANCOM"
author: "Stefano Romano"
output: html_document
---
# Before you begin:

These scripts were tailored for the analyses we conducted. They are not meant to be used in other contexts and are reported here as a detailed extension of the method section of our article. Please, use them at your own risk.

If you use any of the material or approaches in these scripts, please, cite:
Romano et al, 2020, *Meta-analysis of the Parkinson’s disease gut microbiome suggests alterations linked to intestinal inflammation*

Thank you!


```{r setup, include=FALSE}

list.of.packages <- c("dplyr", "reshape2", "dplyr", "propr", "doParallel", "foreach")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

if (!requireNamespace("BiocManager", quietly = TRUE)){
  install.packages("BiocManager")}


list.of.bioc <- c("phyloseq", "zCompositions", "propr", "microbiome")
new.packages <- list.of.bioc[!(list.of.bioc %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)


library(phyloseq)
library(doParallel)
library(foreach)
library(dplyr)

library(microbiome)
library(reshape2)
library(devtools)
library(zCompositions)
library(propr)
source("ancom_v2.0.R")

```

# Overview of the analyses
We will perform a differential abundance analysis using ANCOM2
Please provide the full path to the files and R objects that need to be loaded.

## Load the list pf phyloseq objects
```{r load, message=FALSE}
l_phylo<-readRDS("Phyloseq_list_all_studies.rds")
```

## Remove Eukaryotes and Chloroplast then format data
```{r, message=F}
# Remove Eukaryotes
l_phylo<-lapply(l_phylo, function(x) prune_taxa(taxa_names(x)[!(grepl("Eukaryota", taxa_names(x), ignore.case = TRUE))], x))

# Check that there are not Euk left
if(length(unlist(lapply(l_phylo, function(x) taxa_names(x)[grep("Eukary*", taxa_names(x))]))) == 0){
  print("All Eukaryotes removed")
}else{stop("There are still Eukaryotes sequences in the files")}

# Check Chloroplasts
for(i in 1:length(l_phylo)){
  if(sum(grep("Chloroplast", taxa_names(l_phylo[[i]]))) > 0){
    l_phylo[[i]]<-prune_taxa(taxa_names(l_phylo[[i]])[!(grepl("Chloropl*", taxa_names(l_phylo[[i]]), ignore.case = TRUE))], l_phylo[[i]])
  } else {}
}

if(length(unlist(lapply(l_phylo, function(x) taxa_names(x)[grep("Chloropl*", taxa_names(x))]))) == 0){
  print("All Chloroplast removed")
}else{stop("There are still Chloroplast sequences in the files")}

# Remove samples that have potentially 0 counts
l_phylo<-lapply(l_phylo, function(x) prune_samples(sample_sums(x) > 0, x))

# Format data fro ANCOM (Needs to have easy taxa names)
l_maps<-lapply(l_phylo, function(x) data.frame(Sample.ID = as.character(sample_names(x)), meta(x)))

l_phylo_spec<-l_phylo
for(i in 1:length(l_phylo)){
  taxa_names(l_phylo_spec[[i]]) <- paste0("Species", seq(ntaxa(l_phylo[[i]])))
}

l_spec_taxa<-l_phylo
for(i in 1:length(l_phylo)){
  l_spec_taxa[[i]] <- data.frame(taxa = taxa_names(l_phylo[[i]]), spec = taxa_names(l_phylo_spec[[i]]))
}

l_otu<-lapply(l_phylo_spec, function(x) data.frame(Sample.ID = sample_names(x), t(otu_table(x))))

# Tranform variales into factors
for(i in 1:length(l_maps)){
  l_maps[[i]]$Sample.ID<-as.factor(l_maps[[i]]$Sample.ID)
  l_maps[[i]]$status<-as.factor(l_maps[[i]]$status)
  l_maps[[i]]$study<-as.factor(l_maps[[i]]$study)

}

# ANCOM using parallel
l_anc<-l_maps
cl<-makeCluster(10)
registerDoParallel(cl)
l_anc<-foreach(z = 1:(length(l_otu)-1),
               .packages = c("exactRankTests","coin", "nlme", "ggplot2", "dplyr")) %dopar% {
  x<-ANCOM(otu_data  = l_otu[[z]],
           meta_data  = l_maps[[z]],
           zero_cut = 0.95,
           p_adjust_method = "BH",
           alpha = 0.05,
           main_var = "status",
           rand_formula = NULL,
           adj_formula = NULL)
  return(x)
}
stopCluster(cl)

# Random effect ANCOM for the combined dataset
l_anc[[length(l_otu)]]<-ANCOM(otu_data  = l_otu[[length(l_otu)]],
                              meta_data  = l_maps[[length(l_otu)]],
                              zero_cut = 0.975,
                              p_adjust_method = "BH",
                              alpha = 0.05,
                              main_var = "status",
                              rand_formula = "~ 1 + status|study",
                              adj_formula = NULL)


saveRDS(l_anc, "List_ancom_Species.rds")
saveRDS(l_spec_taxa, "List_Spec_taxa_Species.rds")
```

# Save wrokspace
```{R}
save.image("Species_ANCOM_v2.0_.RData")

```
