---
title: "Combine DA analyses"
author: "Stefano Romano"
output: html_document
  chunk_output_type: console
---
# Before you begin:

These scripts were tailored for the analyses we conducted. They are not meant to be used in other contexts and are reported here as a detailed extension of the method section of our article. Please, use them at your own risk.

If you use any of the material or approaches in these scripts, please, cite:
Romano et al, 2020, *Meta-analysis of the Parkinson’s disease gut microbiome suggests alterations linked to intestinal inflammation*

Thank you!


```{r setup, include=FALSE}

list.of.packages <- c("dplyr", "reshape2", "plyr", "propr", "ggpubr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

if (!requireNamespace("BiocManager", quietly = TRUE)){
  install.packages("BiocManager")}


list.of.bioc <- c("phyloseq", "zCompositions", "microbiome")
new.packages <- list.of.bioc[!(list.of.bioc %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)


library(phyloseq)
library(microbiome)
library(reshape2)
library(dplyr)
library(vegan)
library(devtools)
library(ggpubr)
library(propr)

```

# Overview of the analyses
We combine the data from all different approaches used to infer differential abundance and format dfs for GLMM and for making the final graphs.

Please, provide the full path to the files and R objects that need to be loaded.

## Load the list of objects
```{r}

l_prop<-readRDS("List_prop_species.rds")
x<-readRDS("NP_list_species_DA.rds")
wilcox_prop<-lapply(x, names)
wilcox_prop$Combined<-x$Combined

dseq<-readRDS("DeSeq2_GMPR_list_Species_DA.rds")
rf_ss_ancom<-readRDS("List_ancom_Species_April20.rds")
rf_ss_taxa<-readRDS("List_Spec_taxa_Species_April20.rds")


# Meta-analysis
random.meta.f.all<-readRDS(file = "Meta_species.rds")

# Adjust the results of the meta-analysis.
# Correct the P-values and select the one with sign that have q <= 0.05
random.meta.f.all$pval.random<-p.adjust(random.meta.f.all$pval.random, "BH")
sign<-subset(random.meta.f.all,  (lower.random > 0 & upper.random > 0) | (lower.random < 0 & upper.random < 0))
random.sign.all<-subset(sign, pval.random <= 0.05)

```

## Format ANCOM results
```{r}
ancom_sign_rfss<-lapply(rf_ss_ancom, function(x) subset(x, detected_0.6 == TRUE))
l_new_ancom_sign_taxa<-ancom_sign_rfss

for(i in 1:length(ancom_sign_rfss)){
  l_new_ancom_sign_taxa[[i]]<-merge(ancom_sign_rfss[[i]], rf_ss_taxa[[i]], by.x= 1, by.y=2)$taxa
}

lapply(ancom_sign_rfss, nrow)
l_ancom_sign_taxa<-l_new_ancom_sign_taxa

```

## Make a first graph with number of DA taxa
```{r}
# Get the number for the graph
nw<-lapply(wilcox_prop, function(x) data.frame(nsign=length(x), method = rep("Coin_propor")))
nd<-lapply(dseq, function(x) data.frame(nsign=length(x), method = rep("DeSeq2_GMPR")))
na<-lapply(l_ancom_sign_taxa, function(x) data.frame(nsign=length(x), method = rep("ANCOM")))

da_all<-dseq
all<-dseq
for(i in 1:length(da_all)){
  x<-c(as.character(wilcox_prop[[i]]), as.character(dseq[[i]]), 
       as.character(l_ancom_sign_taxa[[i]]))
  da_all[[i]]<-unlist(unique(x[duplicated(x)])) # this create a list containing only names of DA taxa based on consensus (2 out of 3 approaches)
  all[[i]]<-data.frame(nsign=length(da_all[[i]]), method = "Consensus") # this create a df containing the counts of DA based on consensus (2 out of 3 approaches)
}

l_comb<-all
for(i in 1:length(all)){
  l_comb[[i]]<-rbind.data.frame(nw[[i]], nd[[i]], na[[i]], all[[i]]) # put together in a df the counts of DA recovered by the 3 methods plus the consensus
}
df_da<-data.table::rbindlist(l_comb, use.names = TRUE, idcol = TRUE) %>% as.data.frame() # create a formatted df for plotting

df_da$.id<-factor(df_da$.id, levels = unique(df_da$.id))
saveRDS(df_da, "DA_species.rds")

# Save taxa list of the methods using pooled data
saveRDS(da_all$Combined, "Sign_pooled_data.rds")

# Save data of significant taxa obtained using pooled results (meta_analysis approach)
saveRDS(random.sign.all, "Random_sign_all_species.rds")

taxa_pooled<-unique(c(da_all$Combined, random.sign.all$statistic))

sign_all<-prune_taxa(taxa_pooled, l_prop[[length(l_prop)]])
saveRDS(sign_all, "Phyloseq_species_sign.rds")
```

### Now the external script for calculating GLMM (08) needs to be run

## Save workspace
```{r}
save.image("Combine_DA_species_a.RData")
```

