---
title: "GLMM"
author: "Stefano Romano"
output: html_document
editor_options: 
  chunk_output_type: console
---
# Before you begin:

These scripts were tailored for the analyses we conducted. They are not meant to be used in other contexts and are reported here as a detailed extension of the method section of our article. Please, use them at your own risk.

If you use any of the material or approaches in these scripts, please, cite:
Romano et al, 2020, *Meta-analysis of the Parkinson’s disease gut microbiome suggests alterations linked to intestinal inflammation*

Thank you!

```{r setup, include=FALSE}

list.of.packages <- c("dplyr", "ggplot2", "dplyr", "propr", "ggpubr", "glmmTMB", "bbmle", "MuMIn", "reshape2")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

if (!requireNamespace("BiocManager", quietly = TRUE)){
  install.packages("BiocManager")}


list.of.bioc <- c("phyloseq", "microbiome")
new.packages <- list.of.bioc[!(list.of.bioc %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)

library(glmmTMB)
library(ggplot2)
library(MuMIn)
library(bbmle)
library(phyloseq)
library(microbiome)
library(reshape2)
library(dplyr)
library(ggpubr)
library(propr)

```

# Overview of the analyses
Using the count tables we perform GLMM to verify the effect of gender and age on taxa abundances.
We use the metadata provided by 5 studies.
We calculate GLMMs with and without the variable disease status. 
We then compare all models and select the best ones within a delta AIC of 2.
If among these models there is even a single one that does not contain the variable disease status, we conclude that the disease status is not essential to explain the abundances of these taxa. Hence, we excluded these taxa from further analyses.

Please provide the full path for the files and R objects you need to load.

## Load the meta_data
Meta data table are loaded and formatted to make them consistent to each other

```{r}
hb_meta<-read.delim("HillBurns_Meta.txt", 
                    header = T,
                    stringsAsFactors = F,
                    na.strings = "NA")
hb_meta<-hb_meta[hb_meta$age != "not applicable",]
hb_meta$study<-rep("Hill_Burns", nrow(hb_meta))

any(is.na(hb_meta)) # Check if there are NA

petr_meta<-read.delim("Petrov_meta.txt", 
                    header = T,
                    stringsAsFactors = F,
                    na.strings = "NA")
any(is.na(petr_meta)) # Check if there are NA
petr_meta$study<-rep("Petrov", nrow(petr_meta))

qian_meta<-read.delim("Qian_meta.txt", 
                    header = T,
                    stringsAsFactors = F,
                    na.strings = "NA")
qian_meta<-qian_meta[, 1:6]
qian_meta$gender..1man.2female.[qian_meta$gender..1man.2female. == 1] <- "Male"
qian_meta$gender..1man.2female.[qian_meta$gender..1man.2female. == 2] <- "Female"
qian_meta$gender..1man.2female.[qian_meta$gender..1man.2female. == 3] <- NA
qian_meta<-qian_meta[!(is.na(qian_meta$gender..1man.2female.)),] # remove samples data containing NA
qian_meta$study<-rep("Qian", nrow(qian_meta))

any(is.na(qian_meta)) # Check if there are NA

kes_pd_meta<-read.delim("Kes_meta_PD.txt", 
                    header = T,
                    stringsAsFactors = F,
                    na.strings = "NA")
kes_pd_meta<-data.frame(saples = paste(as.character(kes_pd_meta$PD.Case.IDs), "_136", sep = ""), gender = kes_pd_meta$Gender, age = kes_pd_meta$Age)
kes_pd_meta$study<-rep("Keshavarzan", nrow(kes_pd_meta))

any(is.na(kes_pd_meta)) # Check if there are NA

kes_hc_meta<-read.delim("Kes_meta_HC.txt", 
                    header = T,
                    stringsAsFactors = F,
                    na.strings = "NA")
kes_hc_meta<-data.frame(saples = paste(as.character(kes_hc_meta$HC.Case.IDs), "_136", sep = ""), gender =  as.character(kes_hc_meta$Gender), age =  as.character(kes_hc_meta$Age))
kes_hc_meta$study<-rep("Keshavarzan", nrow(kes_hc_meta))


any(is.na(kes_hc_meta)) # Check if there are NA

weis_meta<-read.delim("Weis_meta.txt", 
                    header = T,
                    stringsAsFactors = F,
                    na.strings = "NA")

weis_meta$study<-rep("Weis", nrow(weis_meta))

any(is.na(weis_meta)) # Check if there are NA

# Select only common confounders: gendere, age

hb_meta$sex[hb_meta$sex == "male"] <-"Male"
hb_meta$sex[hb_meta$sex == "female"] <-"Female"

petr_meta$gender[petr_meta$gender== "f"] <-"Female"
petr_meta$gender[petr_meta$gender== "m"] <-"Male"

weis_meta$gender[weis_meta$gender== "f"] <-"Female"
weis_meta$gender[weis_meta$gender== "m"] <-"Male"

sample = c(as.character(kes_hc_meta$saples), as.character(kes_pd_meta$saples), hb_meta$sample_name, petr_meta$X.SampleID, qian_meta$Sample,
                            weis_meta$sample)
class(sample)
gender = c(as.character(kes_hc_meta$gender), as.character(kes_pd_meta$gender), hb_meta$sex, petr_meta$gender, qian_meta$gender..1man.2female.,
                            weis_meta$gender)
class(gender)
age = c(as.numeric(as.character(kes_hc_meta$age)), as.numeric(as.character(kes_pd_meta$age)), as.numeric(hb_meta$age), petr_meta$Age, qian_meta$age,
                        weis_meta$age)
class(age)

study = c(as.character(kes_hc_meta$study), as.character(kes_pd_meta$study), hb_meta$study, petr_meta$study, qian_meta$study,
                            weis_meta$study)
class(stusy)


conf<-cbind.data.frame(sample = sample, gender = gender, age=age, study =study)
conf$sample<-as.character(conf$sample)
nrow(conf)
any(is.na(conf)) # There are no NA
write.table(x = conf, file = "Table_confounders.txt",
            row.names = F,
            col.names = T,
            sep = "\t",
              quote = F)
```

### Read the Phyloseq object containing all studies and the combined dataset. These data are not rearefied and not normilized

```{r}

l_phylo<-readRDS("Phyloseq_list_all_studies.rds")
sign_otu<-taxa_names(readRDS("Phyloseq_species_sign.rds")) # Load the taxa that were detected DA

# Remove Euk
l_phylo<-lapply(l_phylo, function(x) prune_taxa(taxa_names(x)[!(grepl("Eukaryota", taxa_names(x), ignore.case = TRUE))], x))
# Check that there are not Euk left
if(length(unlist(lapply(l_phylo, function(x) taxa_names(x)[grep("Eukary*", taxa_names(x))]))) == 0){
  print("All Eukaryotes removed")
}else{stop("There are still Eukaryotes sequences in the files")}

# Check Chloroplasts
for(i in 1:length(l_phylo)){
  if(sum(grep("Chloroplast", taxa_names(l_phylo[[i]]))) > 0){
    l_phylo[[i]]<-prune_taxa(taxa_names(l_phylo[[i]])[!(grepl("Chloropl*", taxa_names(l_phylo[[i]]), ignore.case = TRUE))], l_phylo[[i]])
  } else {}
}

if(length(unlist(lapply(l_phylo, function(x) taxa_names(x)[grep("Chloropl*", taxa_names(x))]))) == 0){
  print("All Chloroplast removed")
}else{stop("There are still Chloroplast sequences in the files")}

# Remove samples with 0 counts if any
l_phylo<-lapply(l_phylo, function(x) prune_samples(sample_sums(x) > 0, x))

```

## Select only the samples of interest
There is a bias in the samples, as in some studies the author sampled also control having < 40 years old. Wherease this was not the case for PD patients. 
Hence I will only collect the samples starting form the youngest PD patient.

```{r}
rar_com<-rarefy_even_depth(l_phylo$Combined, sample.size = 10000, replace = FALSE, trimOTUs = TRUE, rngseed = 56987) # rarefy at 10000 to have enough data to limit overparametrization

conf$sample<-as.character(conf$sample)
conf$age<-as.numeric(as.character(conf$age))

conf<-conf[conf$sample %in% sample_names(rar_com),] # use only samples that were included in our meta-analysis. Some other samples might have been used in the original study for other analyses (e.g. microbiome of biopsis)
rar_com_m<-prune_samples(conf$sample, rar_com) # select in the rarefied dataset only the sample of interest
nrow(conf) == nsamples(rar_com_m) # Check that the phyloseq object and the meta data have the same number of samples
rar_com_sign<-prune_taxa(sign_otu, rar_com_m) # select only taxa that were detected DA

# Create a df containing the metadata and the count data
rar_map<-as.data.frame(sample_data(rar_com_sign)) # recover metadata from phyloseq
otu<-as.data.frame(t(otu_table(rar_com_sign))) # recover counts from phyloseq

conf<-conf[order(conf$sample),]
otu<-otu[order(rownames(otu)),]
rar_map<-rar_map[order(rownames(rar_map)),]
all(conf$sample == rownames(otu))
all(rownames(rar_map) == rownames(otu))

# If the above is true means that all meta data, confounders, and tab have the same samples. I can combine them in a single df
data<-cbind.data.frame(otu, rar_map, conf)

# Remove the potential bias due to unbalance sampling
d<-melt(data.frame(samle = data$sample, age = data$age, status = data$status))
ggplot(d, aes(value, fill = status)) +
  geom_bar()

# The first Parkinson case is 43 years old. To avoid further biases I will remove everything < 43
data<-subset(data, age >= 43)
data<-data[,-ncol(data)]
```


## GLMM

```{R}
l_gm_full<-vector(mode = "list", length = (ncol(data)-5)) # create a list to wich I will append the GLMM for each taxa
names(l_gm_full)<-names(data)[1:(ncol(data)-5)]
data$status<-as.factor(data$status)
data$study<-as.factor(data$study)
data2  <- base::transform(data,age=scale(age,center=FALSE)) # rescale the variable. The authors of glmmTMB suggests that this sometimes helps to fit the model

# Fit the full modles including zi inflation
for(i in 1:length(l_gm_full)){
  m<-glmmTMB(data[,i] ~ status * gender + status * age + (1 + status|study), data = data, family = nbinom1, ziformula = ~1,
             control = glmmTMBControl(parallel = 2, optCtrl=list(iter.max=1e3,eval.max=1e3)))
  mcd2<-try(glmmTMB(data2[,i] ~ status * gender + status * age +  (1 + status|study), data = data2, family = nbinom1, ziformula = ~1,
            control = glmmTMBControl(parallel = 2, optCtrl=list(iter.max=1e3,eval.max=1e3))))
  x<-try(summary(mcd2))
  if(class(x) == "try-error"){
    l_gm_full[[i]]<-m
  } else {
    if(!(all(is.na(summary(mcd2)$AICtab[1:4]))) & all(is.na((summary(m)$AICtab[1:4]))) ){
        l_gm_full[[i]]<-mcd2
    } else {l_gm_full[[i]]<-m}
  }
}

sum(unlist(lapply(l_gm_full, function(x) all(is.na((summary(x)$AICtab[1:4]))))))
sum(unlist(lapply(l_gm_full, function(x) all(is.na(summary(x)$coefficients$cond[,2:4])))))
    
# Verify whether there are some other distributions more appropriate for the data used
for(i in 1:length(l_gm_full)){
  if(all(is.na(summary(l_gm_full[[i]])$AICtab[1:4]))){
    m<-update(l_gm_full[[i]], family = nbinom2)
    if(!(all(is.na(summary(m)$AICtab[1:4])))){
      l_gm_full[[i]]<-m
    } else {
      m<-try(update(l_gm_full[[i]], family = genpois))
      if(class(m) == "try-error"){}
      else if(!(all(is.na(summary(m)$AICtab[1:4])))){
        l_gm_full[[i]]<-m
      }
    }
  }
}

sum(unlist(lapply(l_gm_full, function(x) all(is.na((summary(x)$AICtab[1:4]))))))
sum(unlist(lapply(l_gm_full, function(x) all(is.na(summary(x)$coefficients$cond[,2:4])))))

# Fit the full models excluding zi inflation
l_gm_nozero<-l_gm_full
names(l_gm_nozero)<-names(l_gm_full)
for(i in 1:length(l_gm_full)){
  m<-glmmTMB(data[,i] ~ status * gender + status * age +   (1 + status|study), data = data, family = nbinom1, ziformula = ~0,
            control = glmmTMBControl(parallel = 2, optCtrl=list(iter.max=1e3,eval.max=1e3)))
  mcd2<-try(glmmTMB(data2[,i] ~ status * gender + status * age + (1 + status|study), data = data2, family = nbinom1, ziformula = ~0,
            control = glmmTMBControl(parallel = 2, optCtrl=list(iter.max=1e3,eval.max=1e3))))
  x<-try(summary(mcd2))
  if(class(x) == "try-error"){
    l_gm_nozero[[i]]<-m
  } else {
    if(!(all(is.na(summary(mcd2)$AICtab[1:4]))) & all(is.na((summary(m)$AICtab[1:4]))) ){
        l_gm_nozero[[i]]<-mcd2
    } else {l_gm_nozero[[i]]<-m}
  }
}

sum(unlist(lapply(l_gm_nozero, function(x) all(is.na(summary(x)$coefficients$cond[,2:4])))))
sum(unlist(lapply(l_gm_nozero, function(x) all(is.na((summary(x)$AICtab[1:4]))))))

# Verify whether there are some other distributions more appropriate for the data used
for(i in 1:length(l_gm_nozero)){
  if(all(is.na(summary(l_gm_nozero[[i]])$AICtab[1:4]))){
    m<-update(l_gm_nozero[[i]], family = nbinom2)
    if(!(all(is.na(summary(m)$AICtab[1:4])))){
      l_gm_nozero[[i]]<-m
    } else {
      m<-try(update(l_gm_nozero[[i]], family = genpois))
      if(class(m) == "try-error"){}
      else if(!(all(is.na(summary(m)$AICtab[1:4])))){
        l_gm_nozero[[i]]<-m
      }
    }
  }
}


sum(unlist(lapply(l_gm_nozero, function(x) all(is.na(summary(x)$coefficients$cond[,2:4])))))
sum(unlist(lapply(l_gm_nozero, function(x) all(is.na((summary(x)$AICtab[1:4]))))))


# Create a df where I collect the results of the comparison between zi inflated and not-inflated models
aic<-data.frame(taxa = rep(FALSE, length(l_gm_full)),
                           AIC1 = rep(FALSE, length(l_gm_full)),
                           AIC2 = rep(FALSE, length(l_gm_full)),
                           df1 = rep(FALSE, length(l_gm_full)),
                           df2 = rep(FALSE, length(l_gm_full)),
                           DiffAIC = rep(FALSE, length(l_gm_full)),
                           bestmodel = rep(FALSE, length(l_gm_full)))

for(i in 1:length(l_gm_full)){
  if(any(is.na(summary(l_gm_full[[i]])$AICtab)) | any(is.na(summary(l_gm_nozero[[i]])$AICtab)) ){
      aic$taxa[i]<-names(l_gm_full)[i]
      aic$bestmodel[i]<-"No best. Keep PD"
  }else{
    b<-AICtab(l_gm_full[[i]], l_gm_nozero[[i]])
    aic$taxa[i]<-names(l_gm_full)[i]
    aic$AIC1[i]<-b$dAIC[1]
    aic$AIC2[i]<-b$dAIC[2]
    aic$df1[i]<-b$df[1]
    aic$df2[i]<-b$df[2]
    aic$DiffAIC[i]<-b$dAIC[1] - b$dAIC[2]
    aic$bestmodel[i]<-rownames(as.data.frame(b))[1]
  }
}

# Select the best model between zi-inflated and non-inflated models
l_gm_clean2<-l_gm_full
for(i in 1:length(l_gm_full)){
  if(abs(aic$DiffAIC[i]) == 0 & is.na(aic$DiffAIC[i]) == 0 ){
    l_gm_clean2[[i]]<-l_gm_full[[i]]
  } else if(aic[i, "bestmodel"] == "l_gm_nozero[[i]]" & abs(aic$DiffAIC[i]) >= 2 ){
    l_gm_clean2[[i]]<-l_gm_nozero[[i]]
  } else if(aic[i, "bestmodel"] == "l_gm_full[[i]]" & abs(aic$DiffAIC[i]) >= 2  ){ 
    l_gm_clean2[[i]]<-l_gm_nozero[[i]]
  } else if(aic[i, "bestmodel"] == "l_gm_nozero[[i]]" & abs(aic$DiffAIC[i]) < 2){
    l_gm_clean2[[i]]<-l_gm_full[[i]]
  } else if(aic[i, "bestmodel"] == "l_gm_full[[i]]" & abs(aic$DiffAIC[i]) < 2  ){ 
  l_gm_clean2[[i]]<-l_gm_nozero[[i]]
  }
}

sum(unlist(lapply(l_gm_clean2, function(x) all(is.na(summary(x)$coefficients$cond[,2:4])))))
sum(unlist(lapply(l_gm_clean2, function(x) all(is.na((summary(x)$AICtab[1:4]))))))


# Try to select the best model
# Use a backward approach to eliminate each single term one at the time

# No age
l_gm_noage<-l_gm_clean2
for(i in 1:length(l_gm_clean2)){
  l_gm_noage[[i]]<-update(l_gm_clean2[[i]],  ~ status*gender +  (1+status|study))
} 
l_gm_noage_noint<-l_gm_clean2
for(i in 1:length(l_gm_clean2)){
  l_gm_noage_noint[[i]]<-update(l_gm_clean2[[i]],  ~ status+gender +  (1+status|study))
}
l_gm_noage_norandom<-l_gm_clean2
for(i in 1:length(l_gm_clean2)){
  l_gm_noage_norandom[[i]]<-update(l_gm_clean2[[i]],  ~ status*gender +  (1|study))
}
l_gm_noage_norandom_noint<-l_gm_clean2
for(i in 1:length(l_gm_clean2)){
  l_gm_noage_norandom_noint[[i]]<-update(l_gm_clean2[[i]],  ~ status+gender +  (1|study))
}

# No gender
l_gm_nogender<-l_gm_clean2
for(i in 1:length(l_gm_clean2)){
  l_gm_nogender[[i]]<-update(l_gm_clean2[[i]],  ~ status*age +  (1+status|study))
}    
l_gm_nogender_noint<-l_gm_clean2
for(i in 1:length(l_gm_clean2)){
  l_gm_nogender_noint[[i]]<-update(l_gm_clean2[[i]],  ~ status+age +  (1+status|study))
}    
l_gm_nogender_norandom<-l_gm_clean2
for(i in 1:length(l_gm_clean2)){
  l_gm_nogender_norandom[[i]]<-update(l_gm_clean2[[i]],  ~ status*age +  (1|study))
}
l_gm_nogender_norandom_noint<-l_gm_clean2
for(i in 1:length(l_gm_clean2)){
  l_gm_nogender_norandom_noint[[i]]<-update(l_gm_clean2[[i]],  ~ status+age +  (1|study))
}


# Full no interaction
l_gm_noint<-l_gm_clean2
for(i in 1:length(l_gm_clean2)){
  l_gm_noint[[i]]<-update(l_gm_clean2[[i]],  ~ status + age + gender + (1+status|study))
}
l_gm_noint_norandom<-l_gm_clean2
for(i in 1:length(l_gm_clean2)){
  l_gm_noint_norandom[[i]]<-update(l_gm_clean2[[i]],  ~ status + age + gender + (1|study))
}
l_gm_noageint<-l_gm_clean2
for(i in 1:length(l_gm_clean2)){
  l_gm_noageint[[i]]<-update(l_gm_clean2[[i]],  ~ status + age +  status*gender + (1+status|study))
}    
l_gm_noageint_norandom<-l_gm_clean2
for(i in 1:length(l_gm_clean2)){
  l_gm_noageint_norandom[[i]]<-update(l_gm_clean2[[i]],  ~ status + age +  status*gender + (1|study))
}    
l_gm_nogenderint<-l_gm_clean2
for(i in 1:length(l_gm_clean2)){
  l_gm_nogenderint[[i]]<-update(l_gm_clean2[[i]],  ~ status*age + gender + (1+status|study))
}   
l_gm_nogenderint_norandom<-l_gm_clean2
for(i in 1:length(l_gm_clean2)){
  l_gm_nogenderint_norandom[[i]]<-update(l_gm_clean2[[i]],  ~ status*age + gender + (1|study))
}   


# Only status
l_gm_onlystatus<-l_gm_clean2
for(i in 1:length(l_gm_clean2)){
  l_gm_onlystatus[[i]]<-update(l_gm_clean2[[i]],  ~ status + (1+status|study))
} 
l_gm_onlystatus_norandom<-l_gm_clean2
for(i in 1:length(l_gm_clean2)){
  l_gm_onlystatus_norandom[[i]]<-update(l_gm_clean2[[i]], ~ status + (1|study))
}   

# Only age/gender
l_gm_onlygender<-l_gm_clean2
for(i in 1:length(l_gm_clean2)){
  l_gm_onlygender[[i]]<-update(l_gm_clean2[[i]], ~ gender + (1|study))
}    
l_gm_onlyage<-l_gm_clean2
for(i in 1:length(l_gm_clean2)){
  l_gm_onlyage[[i]]<-update(l_gm_clean2[[i]], ~ age + (1|study))
}    
l_gm_nostatus<-l_gm_clean2
for(i in 1:length(l_gm_clean2)){
  l_gm_nostatus[[i]]<-update(l_gm_clean2[[i]],  ~ gender + age + (1|study))
}


# Create a list of models and compare them, collecting the results in a df
model_by_taxa<-vector(mode = "list", length = length(l_gm_clean2))
names(model_by_taxa)<-names(l_gm_clean2)
models<-list(l_gm_clean2, 
             l_gm_nostatus, 
             l_gm_noage, 
             l_gm_nogender, 
             l_gm_noageint, 
             l_gm_nogenderint, 
             l_gm_onlystatus, 
             l_gm_onlystatus_norandom, 
             l_gm_onlygender, 
             l_gm_onlyage, 
             l_gm_noint, 
             l_gm_noint_norandom, 
             l_gm_noage_norandom, 
             l_gm_noage_norandom_noint, 
             l_gm_noage_noint, 
             l_gm_nogender_noint, 
             l_gm_nogender_norandom, 
             l_gm_nogender_norandom_noint,
             l_gm_noageint_norandom, 
             l_gm_nogenderint_norandom)


names(models)<-c("l_gm_clean2",
                 "l_gm_nostatus", 
                 "l_gm_noage", 
                 "l_gm_nogender", 
                 "l_gm_noageint", 
                 "l_gm_nogenderint", 
                 "l_gm_onlystatus", 
                 "l_gm_onlystatus_norandom", 
                 "l_gm_onlygender", 
                 "l_gm_onlyage", 
                 "l_gm_noint", 
                 "l_gm_noint_norandom", 
                 "l_gm_noage_norandom", 
                 "l_gm_noage_norandom_noint", 
                 "l_gm_noage_noint",
                 "l_gm_nogender_noint",
                 "l_gm_nogender_norandom", 
                 "l_gm_nogender_norandom_noint", 
                 "l_gm_noageint_norandom", 
                 "l_gm_nogenderint_norandom")

model_by_taxa<-lapply(model_by_taxa, function(x) x<-vector(mode = "list", length = length(models)))

for(i in 1:length(model_by_taxa)){
  for(m in 1:length(models)){
    x<-try(model.sel(models[[m]][[i]]))
    if(class(x) != "try-error"){
      model_by_taxa[[i]][[m]]<-models[[m]][[i]]
      names(model_by_taxa[[i]])[m]<-names(models)[m]
    }
  }
  model_by_taxa[[i]]<-model_by_taxa[[i]][!(is.na(names(model_by_taxa[[i]])))]
}



l_mod_sel2<-model_by_taxa
for(i in 1:length(l_mod_sel2)){
  l_mod_sel2[[i]]<-model.sel(model_by_taxa[[i]],
                             rank =  AIC)
}

```

## Detect the taxa that have abundances mainly affected by age and/or gender

```{R}

# Take a look at the results of the comparisons and verify whether there is a single best model that explains the data
for(i in 1:length(l_mod_sel2)){
  l_mod_sel2[[i]]$taxa<-rep(names(l_mod_sel2)[[i]], nrow(l_mod_sel2[[i]]))
  l_mod_sel2[[i]]$models<-rownames(as.data.frame(l_mod_sel2[[i]]))

}

tab<-do.call(rbind.fill, l_mod_sel2)
tab<-cbind.data.frame(taxa = tab$taxa, models = tab$models, tab[-1*c(which(names(tab) == "taxa"), which(names(tab) == "models"))])

# Save the table containing the model comparisons
write.table(tab, file = "GLMM_species.txt", 
            sep = "\t", 
            quote = F, 
            row.names = F, 
            na = "NA", 
            col.names = T)

confounded<-vector()
taxa_w_more_model<-vector()
for(i in 1:length(l_mod_sel2)){
  x<-l_mod_sel2[[i]]
  if( all(is.na(x$delta))){
    print(paste("Model not estimated for taxa", names(l_mod_sel2)[i], sep = " "))
    print(" ", quote = FALSE)
    print(" ", quote = FALSE)
    } else {
      models<-rownames(subset(as.data.frame(x), delta <= 2))
      if(all(models == "l_gm_onlygender" | models == "l_gm_nostatus" | models == "l_gm_onlyage")){
      confounded<-c(confounded, names(l_mod_sel2)[[i]])
      }
    }
}


if(length(confounded) == 0){
  print("No taxa are directly confounder by age/gender. You need to investigate the model evaluation manually")
}else{print(paste("List of taxa confounded", confounded, sep = " "))}

#Save RDS
saveRDS(confounded, "Confounded_species.rds")
```

## Save workspace

```{r}
save.image("Confounders_species.RData")
```