---
title: "Combine graphs"
author: "Stefano Romano"
output: html_document
editor_options: 
  chunk_output_type: console
---
# Before you begin:

These scripts were tailored for the analyses we conducted. They are not meant to be used in other contexts and are reported here as a detailed extension of the method section of our article. Please, use them at your own risk.

If you use any of the material or approaches in these scripts, please, cite:
Romano et al, 2020, *Meta-analysis of the Parkinson’s disease gut microbiome suggests alterations linked to intestinal inflammation*

Thank you!

# Load Packages

```{r setup, include=FALSE}
list.of.packages <- c("ggplot2", "dplyr", "reshape2", "dplyr", "devtools", "ggpubr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

if (!requireNamespace("BiocManager", quietly = TRUE)){
  install.packages("BiocManager")}


list.of.bioc <- c("phyloseq", "microbiome")
new.packages <- list.of.bioc[!(list.of.bioc %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)


library(phyloseq)
library(microbiome)
library(ggpubr)
library(ggplot2)
library(reshape2)
library(dplyr)
library(grDevices)

```

# Overview of the analyses
This script was used to create some of the figures in the manuscript.

Please, provide the full path to the files and R objects that need to be loaded.

# Combine dbRDA graphs for TSS
```{r, message= F}
g_species<-readRDS("dbRDA_species_propr.rds")
g_genus<-readRDS("dbRDA_genus_propr.rds")
g_family<-readRDS("dbRDA_family_propr.rds")

g_propr<-gridExtra::grid.arrange(g_family, g_genus, g_species, nrow = 3, heights = c(1.2,1,1),
                        top = "dbRDA-JSD on TSS")
ggsave(g_propr, filename = "dbRDA_prop.svg",
       width = 12, height = 12, device = "svg", dpi = 300)
```

# Combine graphs for VST
```{r, message= F}
g_species<-readRDS("dbRDA_species_VST.rds")
g_genus<-readRDS("dbRDA_genus_VST.rds")
g_family<-readRDS("dbRDA_family_VST.rds")

g_propr<-gridExtra::grid.arrange(g_family, g_genus, g_species, nrow = 3, heights = c(1.2,1,1),
                                 top = "dbRDA-Euclidean on VST")
ggsave(g_propr, filename = "dbRDA_VST.svg",
       width = 12, height = 12, device = "svg", dpi = 300)
```

# Combine graphs for CLR
```{r, message= F}
g_species<-readRDS("dbRDA_species_CLR.rds")
g_genus<-readRDS("dbRDA_genus_CLR.rds")
g_family<-readRDS("dbRDA_family_CLR.rds")

g_propr<-gridExtra::grid.arrange(g_family, g_genus, g_species, nrow = 3, heights = c(1.2,1,1),
                                 top = "dbRDA-Euclidean on CLR")
ggsave(g_propr, filename = "dbRDA_CLR.svg",
       width = 12, height = 12, device = "svg", dpi = 300)
```

## Plot overall DA taxa
```{r}
# Load files
da_f<-readRDS("DA_Family.rds")
da_g<-readRDS("DA_genus.rds")
da_s<-readRDS("DA_species.rds")
random_f<-readRDS("Random_sign_all_family.rds")
random_g<-readRDS("Random_sign_all_genus.rds")
random_s<-readRDS("Random_sign_all_species.rds")

# Format df and make graphs

da_s$.id<-da_s$.id %>% as.character()
da_s$method<-da_s$method %>% as.character()

da_s<-rbind.data.frame(da_s, c(F, nrow(random_s), F))
da_s$.id[nrow(da_s)]<-"Pooled_results"
da_s$method[nrow(da_s)]<-"Random effect meta."

da_s$.id[da_s$.id == "Combined"] <- "Pooled_data"
da_s$method[da_s$method == "Coin_propor"] <- "WMW_TSS"
da_s$method[da_s$method == "DeSeq2_GMPR"] <- "DESeq"

da_s$method<-factor(da_s$method, levels = c("Random effect meta.", "WMW_TSS","DESeq","ANCOM","Consensus"))

da_s$.id<-factor(da_s$.id, levels = unique(da_s$.id))
s<-ggplot(da_s, aes(nsign, .id, fill = method)) +
  geom_point(pch = 21, alpha = 0.7, size = 4.5) +
  theme_bw() +
  labs(x = "#DA taxa", fill = "Method") +
  scale_fill_brewer(type = "div", palette = 7) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 16),
        legend.text = element_text(size =14),
        legend.title = element_text(size = 14, face = "bold"),
        plot.margin = margin(15,2,2,2),
        axis.text.x = element_text(size = 14))


da_g$.id<-da_g$.id %>% as.character()
da_g$method<-da_g$method %>% as.character()

da_g<-rbind.data.frame(da_g, c(F, nrow(random_g), F))
da_g$.id[nrow(da_g)]<-"Pooled_results"
da_g$method[nrow(da_g)]<-"Random effect meta."

da_g$.id[da_g$.id == "Combined"] <- "Pooled_data"
da_g$method[da_g$method == "Coin_propor"] <- "WMW_TSS"
da_g$method[da_g$method == "DeSeq2_GMPR"] <- "DESeq"

da_g$method<-factor(da_g$method, levels = c("Random effect meta.", "WMW_TSS","DESeq","ANCOM","Consensus"))
da_g$.id<-factor(da_g$.id, levels = unique(da_g$.id))

g<-ggplot(da_g, aes(nsign, .id, fill = method)) +
  geom_point(pch = 21, alpha = 0.7, size = 4.5) +
  theme_bw() +
  labs(x = "#DA taxa", fill = "Method") +
  scale_fill_brewer(type = "div", palette = 7) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 16),
        legend.text = element_text(size =14),
        legend.title = element_text(size = 14, face = "bold"),
        plot.margin = margin(15,2,2,2),
        axis.text.x = element_text(size = 14))


da_f$.id<-da_f$.id %>% as.character()
da_f$method<-da_f$method %>% as.character()

da_f<-rbind.data.frame(da_f, c(F, nrow(random_f), F))
da_f$.id[nrow(da_f)]<-"Pooled_results"
da_f$method[nrow(da_f)]<-"Random effect meta."

da_f$.id[da_f$.id == "Combined"] <- "Pooled_data"
da_f$method[da_f$method == "Coin_propor"] <- "WMW_TSS"
da_f$method[da_f$method == "DeSeq2_GMPR"] <- "DESeq"

da_f$method<-factor(da_f$method, levels = c("Random effect meta.", "WMW_TSS","DESeq","ANCOM","Consensus"))
da_f$.id<-factor(da_f$.id, levels = unique(da_f$.id))

f<-ggplot(da_f, aes(nsign, .id, fill = method)) +
  geom_point(pch = 21, alpha = 0.7, size = 4.5) +
  theme_bw() +
  labs(x = "#DA taxa", fill = "Method", y = "Study") +
  scale_fill_brewer(type = "div", palette = 7) +
  theme(axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size =14),
        legend.title = element_text(size = 14, face = "bold"),
        plot.margin = margin(15,2,2,2),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14))

# Combine graphs and save

g_c<-ggarrange(f,g,s, common.legend = T, 
                font.label = list(size = 12), label.x = 0.15, label.y = 0.975, ncol = 3,
                labels = c("a: Family", "b: Genus", "c: Species"),
                hjust = c(-1.3,0.7,0.6), 
                vjust = c(-0.1,-0.1,-0.1),
                widths = c(1.35,1,1),
                align = "h") 

title1=text_grob( "Number of differentially abundant taxa", size = 18, face = "bold")

g_cf<-annotate_figure(g_c, top = title1)

ggsave(g_cf, 
       filename = "/home/stefano/Desktop/Lotus_PD/07_R_Markdown_9_Sept_2020_noChloro_revision_all10/04_05_Figures/DA_all.tiff",
       device = "tiff", 
       width = 13, 
       height = 6, 
       dpi = 300)

```

## Combine Adonis plots
```{r}
# Load files
prop<-readRDS("Adonis_species_prop.rds")
vst<-readRDS("Adonis_species_VST.rds")
clr<-readRDS("Adonis_species_CLR.rds")


prop_g<-readRDS("Adonis_genus_prop.rds")
vst_g<-readRDS("Adonis_genus_VST.rds")
clr_g<-readRDS("dbRDA_genus_CLR.rds")

prop_f<-readRDS("Adonis_family_prop.rds")
vst_f<-readRDS("Adonis_family_VST.rds")
clr_f<-readRDS("Adonis_family_CLR.rds")

# Format dfs and make graphs
all<-rbind.data.frame(prop, vst,clr)
all$Study<-factor(all$Study, levels = unique(all$Study))
all$Significant<-as.character(all$Significant)
all$Significant<-factor(all$Significant, levels = c("Not_significant","Significant [0.05<p<0.1]","Significant [p<0.05]"))
all$Index[all$Index == "Bray_prop"]<-"BC_TSS"
all$Index[all$Index == "Bray_VST"]<-"BC_VST"
all$Index[all$Index == "JSD_prop"]<-"JSD_TSS"

s<-ggplot(all, aes(R2, Study, color = Index)) +
  geom_point(colour = "black", size = 4.8, aes(shape = Significant)) +
  geom_point(size = 4, aes(shape = Significant)) +
  theme_bw() +
  labs(shape = "Significance") +
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(), 
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 14),legend.text.align = 0, legend.text = element_text(size = 12),
        legend.title = element_text(size = 14, face = "bold"), plot.margin = unit(c(7,2,2,2), units = "mm")) + 
  scale_x_continuous(limits = c(0,8), breaks = c(0,1,2,3,4,5,6,7,8,9)) +
  scale_shape_manual(values=c(15,16,17)) +
  scale_color_viridis_d()

all_g<-rbind.data.frame(prop_g, vst_g, clr_g)
all_g$Study<-factor(all_g$Study, levels = unique(all_g$Study))
all_g$Significant<-as.character(all_g$Significant)
all_g$Significant<-factor(all_g$Significant, levels = c("Not_significant","Significant [0.05<p<0.1]","Significant [p<0.05]"))

all_g$Index[all_g$Index == "Bray_prop"]<-"BC_TSS"
all_g$Index[all_g$Index == "Bray_VST"]<-"BC_VST"
all_g$Index[all_g$Index == "JSD_prop"]<-"JSD_TSS"

g<-ggplot(all_g, aes(R2, Study, color = Index)) +
  geom_point(colour = "black", size = 4.8, aes(shape = Significant)) +
  geom_point(size = 4, aes(shape = Significant)) +
  theme_bw() +
  labs(shape = "Significance") +
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(), axis.title.x = element_text(size =14),
        axis.text.x = element_text(size = 14),legend.text.align = 0, legend.text = element_text(size = 12),
        legend.title = element_text(size = 14, face = "bold"), plot.margin = unit(c(7,2,2,2), units = "mm")) + 
  scale_x_continuous(breaks = c(0,1,2,3,4,5,6,7,8,9)) +
    scale_shape_manual(values=c(15,16,17)) +
    scale_color_viridis_d() +
  xlab(expression(Proportion~of~variance~R^{"2"}))

all_f<-rbind.data.frame(prop_f, vst_f, clr_f)
all_f$Study<-factor(all_f$Study, levels = unique(all_f$Study))
all_f$Significant<-as.character(all_f$Significant)
all_f$Significant<-factor(all_f$Significant, levels = c("Not_significant","Significant [0.05<p<0.1]","Significant [p<0.05]"))

all_f$Index[all_f$Index == "Bray_prop"]<-"BC_TSS"
all_f$Index[all_f$Index == "Bray_VST"]<-"BC_VST"
all_f$Index[all_f$Index == "JSD_prop"]<-"JSD_TSS"

f<-ggplot(all_f, aes(R2, Study, color = Index)) +
  geom_point(colour = "black", size = 4.8, aes(shape = Significant)) +
  geom_point(size = 4, aes(shape = Significant)) +
  theme_bw() +
  labs(shape = "Significance") +
  ylab("Study") +
  theme(legend.position = "none",
        axis.title.x = element_blank(), axis.text.y = element_text(size = 14),axis.title.y = element_text(size = 14),
        axis.text.x = element_text(size = 14),legend.text.align = 0, legend.text = element_text(size = 12),
        legend.title = element_text(size = 14, face = "bold"), plot.margin = unit(c(7,2,2,2), units = "mm")) + 
  scale_x_continuous(limits = c(0,13), breaks = c(0,1,2,3,4,5,6,7,8,9,10,11,12,13,14)) +
  scale_shape_manual(values=c(15,16,17)) +
  scale_color_viridis_d() 

# Combine graphs and save
(g_c<-ggarrange(f,g,s, common.legend = T, align = "h",
               font.label = list(size = 14.5), label.x = 0.85, label.y = 0.975, ncol = 3,
             labels = c("a: Family", "b: Genus", "c: Species"), 
             hjust  = c(3.25,3.4,3), 
             vjust = c(0.5,0.5,0.5),
             widths = c(1.25,0.85,0.85)))
(g_cf<-annotate_figure(g_c, top = text_grob("PERMANOVA: Proportion of variance explained by disease status", size = 16, face = "bold")))


ggsave(g_cf, filename = "ADONIS.svg",
       device = "svg", 
       width = 13, 
       height = 6, 
       dpi = 300)
```

# Combine comparison with and without Scheperjans
```{R}
#Picrust
load("Combine_DA_metacyc.RData")
taxa_wschep_p<-random_only_sign_store
load("Combine_DA_metacyc_noSchep.RData")
taxa_no_schep_p<-random_only_sign
uneven<-c(setdiff(taxa_wschep_p$statistic, taxa_no_schep_p$statistic), setdiff(taxa_no_schep_p$statistic, taxa_wschep_p$statistic))
taxa_wschep_p<-subset(taxa_wschep_p, statistic %in% uneven)
taxa_no_schep_p<-subset(taxa_no_schep_p, statistic %in% uneven)

taxa_wschep_p$schep<-rep("Scheperjans included", nrow(taxa_wschep_p))
taxa_no_schep_p$schep<-rep("Scheperjans excluded", nrow(taxa_no_schep_p))

dfp<-rbind.data.frame(taxa_wschep_p, taxa_no_schep_p)

# Make graphs
psp<-ggplot(dfp, aes(y=TE.random, 
                        ymax=upper.random, 
                        ymin=lower.random,
                        x=reorder(statistic, TE.random))) + 
  geom_errorbar(width=0.2) + 
  geom_point(aes(fill = seTE.random), size = 3.5, pch = 21) +  
  coord_flip() + 
  geom_hline(yintercept=0) + theme_bw() + labs(y="Mean difference in CLR (95% CI)") +
  facet_wrap(~schep) +
  xlab("Metacyc Pathways") +
  scale_fill_viridis_c() +
  labs(fill = "SE")+
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        strip.text.x = element_text(size = 12))

# Family
load("Combine_DA_family_b.RData")
taxa_wschep_f<-random.sign.all
load("Combine_DA_family_b_noSchep.RData")
taxa_no_schep_f<-random.sign.all
uneven_f<-c(setdiff(taxa_wschep_f$statistic, taxa_no_schep_f$statistic), setdiff(taxa_no_schep_f$statistic, taxa_wschep_f$statistic))

taxa_wschep_f<-subset(taxa_wschep_f, statistic %in% uneven_f)
taxa_no_schep_f<-subset(taxa_no_schep_f, statistic %in% uneven_f)

taxa_wschep_f$schep<-rep("Scheperjans included", nrow(taxa_wschep_f))
taxa_no_schep_f$schep<-rep("Scheperjans excluded", nrow(taxa_no_schep_f))
df_f<-rbind.data.frame(taxa_wschep_f, taxa_no_schep_f)

# Make graph
fsp<-ggplot(df_f, aes(y=TE.random, 
                        ymax=upper.random, 
                        ymin=lower.random,
                        x=reorder(statistic, TE.random))) + 
  geom_errorbar(width=0.2) + 
  geom_point(aes(fill = seTE.random), size = 4, pch = 21) +  
  coord_flip() + 
  geom_hline(yintercept=0) + theme_bw() + labs(y="Mean difference in CLR (95% CI)") +
  facet_wrap(~schep) +
  xlab("Taxa: Family") +
  scale_fill_viridis_c() +
  labs(fill = "SE") +
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        strip.text.x = element_text(size = 12))
# Genus

load("Combine_DA_genus_b.RData")
taxa_wschep_g<-random.sign.all
load("Combine_DA_genus_b_noSchep.RData")
taxa_no_schep_g<-random.sign.all
uneven_g<-c(setdiff(taxa_wschep_g$statistic, taxa_no_schep_g$statistic), setdiff(taxa_no_schep_g$statistic, taxa_wschep_g$statistic))
taxa_wschep_g<-subset(taxa_wschep_g, statistic %in% uneven_g)
taxa_no_schep_g<-subset(taxa_no_schep_g, statistic %in% uneven_g)

taxa_wschep_g$schep<-rep("Scheperjans included", nrow(taxa_wschep_g))
taxa_no_schep_g$schep<-rep("Scheperjans excluded", nrow(taxa_no_schep_g))

dfg<-rbind.data.frame(taxa_wschep_g, taxa_no_schep_g)

# Make graph
gsp<-ggplot(dfg, aes(y=TE.random, 
                        ymax=upper.random, 
                        ymin=lower.random,
                        x=reorder(statistic, TE.random))) + 
  geom_errorbar(width=0.2) + 
  geom_point(aes(fill = seTE.random), size = 4, pch = 21) +  
  coord_flip() + 
  geom_hline(yintercept=0) + theme_bw() + labs(y="Mean difference in CLR (95% CI)") +
  facet_wrap(~schep) +
  xlab("Taxa: Genus") +
  scale_fill_viridis_c() +
  labs(fill = "SE") +
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        strip.text.x = element_text(size = 12))


# Species
load("Combine_DA_species_b.RData")
taxa_wschep_s<-random.sign.all
load("Combine_DA_species_b_noSchep.RData")
taxa_no_schep_s<-random.sign.all
uneven_s<-c(setdiff(taxa_wschep_s$statistic, taxa_no_schep_s$statistic), setdiff(taxa_no_schep_s$statistic, taxa_wschep_s$statistic))
taxa_wschep_s<-subset(taxa_wschep_s, statistic %in% uneven_s)
taxa_no_schep_s<-subset(taxa_no_schep_s, statistic %in% uneven_s)

taxa_wschep_s$schep<-rep("Scheperjans included", nrow(taxa_wschep_s))
taxa_no_schep_s$schep<-rep("Scheperjans excluded", nrow(taxa_no_schep_s))

dfs<-rbind.data.frame(taxa_wschep_s, taxa_no_schep_s)

# Make graph
ssp<-ggplot(dfs, aes(y=TE.random, 
                        ymax=upper.random, 
                        ymin=lower.random,
                        x=reorder(statistic, TE.random))) + 
  geom_errorbar(width=0.2) + 
  geom_point(aes(fill = seTE.random), size = 4, pch = 21) +  
  coord_flip() + 
  geom_hline(yintercept=0) + theme_bw() + labs(y="Mean difference in CLR (95% CI)") +
  facet_wrap(~schep) +
  xlab("Taxa: Species") +
  scale_fill_viridis_c() +
  labs(fill = "SE") +
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        strip.text.x = element_text(size = 12))

# Combine graphs and save
gg<-ggarrange(gsp, ssp, psp, ncol = 1, align = "v")
g_cf<-annotate_figure(gg, top = text_grob("Sensitivity analysis performed omitting the dataset from Scheperjans et al.", face  = "bold", size =14))

ggsave(filename = "Diff_sch_overall.tiff", 
       plot = g_cf, 
       device = "tiff",
       width = 11, 
       height = 20,
       dpi = 300)
```

#Prevotella figures across gender and studies

Re-run the part included in the GLMM script to collect only the data of the datasets reporting metadata.
After that, make the graph.

```{R}
# Family
load("Confounders_family.RData")
pro<- transform_sample_counts(rar_com, function(x){x / sum(x)})

conf$sample<-as.character(conf$sample)
conf$age<-as.numeric(as.character(conf$age))

conf<-conf[conf$sample %in% sample_names(pro),]
rar_com_m<-prune_samples(conf$sample, pro)
nrow(conf) == nsamples(rar_com_m) 
rar_com_sign<-prune_taxa(sign_otu, rar_com_m)

# Createa df containing the metadata and the count data
rar_map<-as.data.frame(sample_data(rar_com_sign)) # recover metadata from phyloseq
otu<-as.data.frame(t(otu_table(rar_com_sign))) # recover taxa counts from phyloseq

conf<-conf[order(conf$sample),]
otu<-otu[order(rownames(otu)),]
rar_map<-rar_map[order(rownames(rar_map)),]
all(conf$sample == rownames(otu))
all(rownames(rar_map) == rownames(otu))
data<-cbind.data.frame(otu, rar_map, conf)

# Remove the potential bias due to unbalanced sampling
d<-melt(data.frame(samle = data$sample, age = data$age, status = data$status))
ggplot(d, aes(value, fill = status)) +
  geom_bar()
data<-subset(data, age >= 43)

# Do graph for Prevotellaceae
prev<-cbind.data.frame( rel_ab = data$`Bacteria;Bacteroidota;Bacteroidia;Bacteroidales;Prevotellaceae`,
                        age =  data$age, gender = data$gender,
                        status = data$status,
                        study = data$study)

prev$status<-as.factor(prev$status)
prev$gender<-as.factor(prev$gender)

fam<-ggplot(prev, aes(status, log10(rel_ab), fill = gender)) +
  geom_boxplot() +
  scale_fill_brewer(type = "div", palette = 7) +
  theme_bw() +
  facet_wrap(~ study, ncol = 5)+
  xlab("Status") +
  labs(fill = "Gender") +
  theme(strip.text.x = element_text(size = 12),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 14),
        legend.title = element_text(size =12, face = "bold"),
        axis.text.x = element_text(size = 14),
        legend.text = element_text(size = 12),
        axis.text.y = element_text(size = 14),
        plot.margin = margin(15,2,5,2))

# Species 
load("Confounders_species.RData")
pro<- transform_sample_counts(rar_com, function(x){x / sum(x)})

conf$sample<-as.character(conf$sample)
conf$age<-as.numeric(as.character(conf$age))

conf<-conf[conf$sample %in% sample_names(pro),] 
rar_com_m<-prune_samples(conf$sample, pro) 
nrow(conf) == nsamples(rar_com_m)
rar_com_sign<-prune_taxa(sign_otu, rar_com_m)

rar_map<-as.data.frame(sample_data(rar_com_sign)) # recover metadata from phyloseq
otu<-as.data.frame(t(otu_table(rar_com_sign))) # recover taxa counts from phyloseq

conf<-conf[order(conf$sample),]
otu<-otu[order(rownames(otu)),]
rar_map<-rar_map[order(rownames(rar_map)),]
all(conf$sample == rownames(otu))
all(rownames(rar_map) == rownames(otu))
data<-cbind.data.frame(otu, rar_map, conf)

# Remove the potential bias due to unbalance sampling
d<-melt(data.frame(samle = data$sample, age = data$age, status = data$status))
ggplot(d, aes(value, fill = status)) +
  geom_bar()
data<-subset(data, age >= 43)

# Do graph for Prevotellaceae sp.
prev<-cbind.data.frame( rel_ab = data$`Bacteria;Bacteroidota;Bacteroidia;Bacteroidales;Prevotellaceae;?;?`,
                        age =  data$age, gender = data$gender,
                        status = data$status,
                        study = data$study)

prev$status<-as.factor(prev$status)
prev$gender<-as.factor(prev$gender)

sp<-ggplot(prev, aes(status, log10(rel_ab), fill = gender)) +
  geom_boxplot() +
  scale_fill_brewer(type = "div", palette = 7) +
  theme_bw() +
  facet_wrap(~ study, ncol = 5)+
  xlab("Status") +
  labs(fill = "Gender")+
  theme(strip.text.x = element_text(size = 12),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 14),
        legend.title = element_text(size =12, face = "bold"),
        legend.text = element_text(size = 12),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        plot.margin = margin(15,2,5,2))

# Do graph for Prevotella sp (from gut metagenome)
prev2<-cbind.data.frame( rel_ab = data$`Bacteria;Bacteroidota;Bacteroidia;Bacteroidales;Prevotellaceae;Prevotella;metagenome`,
                        age =  data$age, gender = data$gender,
                        status = data$status,
                        study = data$study)

prev2$status<-as.factor(prev$status)
prev2$gender<-as.factor(prev$gender)

sp_2<-ggplot(prev2, aes(status, log10(rel_ab), fill = gender)) +
  geom_boxplot() +
  scale_fill_brewer(type = "div", palette = 7) +
  theme_bw() +
  facet_wrap(~ study, ncol = 5)+
  xlab("Status") +
  labs(fill = "Gender") +
  theme(strip.text.x = element_text(size = 12),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        legend.title = element_text(size =12, face = "bold"),
        axis.text.x = element_text(size = 14),
        legend.text = element_text(size = 12),
        axis.text.y = element_text(size = 14),
        plot.margin = margin(15,2,5,2))

# Combine graphs
gg<-ggarrange(fam, sp, sp_2, common.legend = T, ncol = 1, align = "v",
          labels = c("a: Family Prevotellaceae", "b: Prevotellaceae unknown sp.", "c: Prevotella sp"),
          hjust = c(-0.30,-0.35, -0.5),
          vjust = c(-0.2,-0.2, -0.2),
          font.label = list(size = 12))

# Add title
g_cf<-annotate_figure(gg, top = text_grob("Distribution of Prevetella across study, gender, and disease status"))

# Save
ggsave(g_cf, filename = "Prev_gender.svg",
       device = "svg", width = 16, height = 4.5, dpi = 300)

```

